---
title: "R Notebook"
output: html_notebook
---

# Unit Testing

```{r}
is_leap_year <- function(year){
  
# add a condition to make sure our function only takes numeric input
  
  if(!is.numeric(year)) {
    stop("Non-numeric input")
  }
  
  if (year %% 100 == 0){
    if (year %% 400 == 0){
      return(TRUE)
    } else{
      return(FALSE)
    }
  } 
  
  if (year %% 4 == 0){
    return(TRUE)
  } else {
    return(FALSE)
  }
}
```

# Practice
Write out what this function is doing, and why it is written in this format (e.g. Why do we test the 100 and 400 years first?)

This function first tests if an input year is divisible by 100 (if year / 100 has no remainder).
If it is divisible by 100, then it checks if the year is divisible by 400.
If the year is divisible by 100 and 400, it returns TRUE.
If the year is divisible by 100, but not 400, it returns FALSE.

It then tests whether the year is divisible by 4 and returns TRUE if it is and FALSE if not.

It checks for divisible by 100 and 400 first because 100 is divisible by 4, but it is not a leap year, so it is weeding out the centenary years which are not leap years.

#The Test That Library

```{r}
library(testthat)
```

## Expect_equal

At its simplest. If it works, nothing happens. 
```{r}
a <- 11
expect_equal(a, 10)
```

```{r}
test_that("Real leap years return true values", {
  expect_equal(is_leap_year(2020), TRUE)
  expect_equal(is_leap_year(2000), TRUE)
  expect_equal(is_leap_year(1900), FALSE)
}
)
```


## Expect error

```{r}
test_that("Character input returns an error", {
  expect_error(is_leap_year("year"))
})
```

## A failed test
Which of our tests has failed? expect error TRUE
Why has it failed? It returned FALSE instead of an error.
It's showing us that our function can run on a Boolean, so we edit our function to only accept numeric values.
Why has our other test passed? Because it returns an error (as expected)

```{r}
test_that("Non-numeric input returns an error", {
  expect_error(is_leap_year("year"))
  expect_error(is_leap_year(TRUE))
})
```

```{r}
is_leap_year(TRUE)
```

# Practice

Write a test that checks whether 2019 and 1900 are correctly identified as not being leap years by our is_leap_year() function.

```{r}
test_that("non-leap years return false value", {
          expect_equal(is_leap_year(2019), FALSE)
          expect_equal(is_leap_year(1900), FALSE)
  }
)
```

