---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Strings basics

```{r}
string <- "a contraction in double quotes isn't a problem"

problem_string <- 'a contraction in double quotes isn't a problem'
```


```{r}
# literally empty
empty_string <- ""

# looks empty
not_an_empty_string <- "   "

# showing that not_an_empty string is considered a string
# This is a common problem when dealing with excel data!
empty_string == not_an_empty_string
# Returns FALSE
```

# Manipulating Strings

## Concatenate

```{r}
str_c("This", "is", "a", "string", sep = " ")
# "This is a string"

default <- str_c("This", "is", "a", "string")
# "Thisisastring"

str_c("This", "is", "a", "string", sep = "_")
# This_is_a_string
```

### Naming files example
```{r}
clinic_id <- "12345_edin"
date <- "2020_12_19"
filename <- str_c(clinic_id, date, sep = "_")
# 12345_edin_2020_12_19
```

### Handling missing values
```{r}
str_c("See", NULL, "how", NULL, "these", NULL, "disappear")
# See how these disappear
```

## Casing

```{r}
pangram <- "The quick brown fox jumps over the lazy dog"

#Upcasing
pangram %>% 
  str_to_upper()

#Lowcasing
pangram %>% 
  str_to_lower()

# Title-ising every word starts with a capital letter
pangram %>% 
  str_to_title()

# Turn string into a sentence
pangram %>% 
  str_to_sentence()
```

## Splitting Strings
Helps us to create vectors of individual words

```{r}
unbroken_string <- "We're going to split this string up"
```

Makes a list
```{r}
broken_string_list <- unbroken_string %>% 
  str_split(" ")
```

Make a vector
```{r}
broken_string_vector <- unbroken_string %>% 
  str_split(" ") %>% 
  flatten_chr()
```

Split on a different separator
```{r}
 unbroken_string %>% 
  str_split("i") %>% 
  flatten_chr()
```

Split seperator default
```{r}
unbroken_string %>% 
  str_split("") %>% 
  flatten_chr()
```

# Practice
Concatenate the strings "Manipulating" "strings" "is" "fun!"
Capitalise the result and save it in a new variable
Split the formatted string into individual characters

```{r}
my_string <- str_c("Manipulating", "strings", "is", "fun!", sep =" ") %>% 
  str_to_upper()
  
my_string %>% 
  str_split("") %>% 
  flatten_chr()
```

## Finding substrings

```{r}
full_string <- "Let's grab a substring from this!"
full_string
```


```{r}
# First five
full_string %>% 
  str_sub(start = 1, end = 5)
```


```{r}
#Last five
full_string %>% 
  str_sub(start = -5, end = -1)
```

```{r}
# using defaults to start at a position and run to the end. Ignores a constant value in a set of strings
full_string %>% 
  str_sub(6)
# reads (start = 6, end = default(-1))
```

```{r}
full_string %>% 
  str_sub(1:5, 10)
# ? I think this works like str_sub((c(1,2,3,4,5), 10)
```

```{r}
full_string %>% 
  str_sub(1, 5:10)
```

```{r}
full_string %>% 
  str_sub(1:5, 6:10)
```

# Replacing substrings

```{r}
original_string <- "We're going to replace some words."

original_string
```

```{r}
str_sub(original_string, 1, 22) <- "I've changed"

original_string
```
```{r}
str_sub(original_string, -11) <- "everything!"

original_string
```

# Numbers in strings

```{r}
number <- 123456.789

#Changed to a string and by default rounds to one decimal place
number %>% 
  format()

#Changing minimum digits to the right of decimal point big = right hand side, small = left hand side
number %>% 
  format(nsmall = 2)

# Changing formatting e.g. decimal mark
number %>% 
  format(nsmall = 2, decimal.mark = ",")

#Changing format e.g commas
number %>% 
  format(big.mark = ",") 


```

# Practice
Concatenate the strings "I" "love" "using" "strings" "in" "reports!"
Replace the substring "I love using" with "Isn't working with"
Replace the substring "in reports!" with "really interesting.
```{r}
string <- str_c("I", "love", "using", "strings", "in", "reports!", sep = " ")

str_sub(string, 1, 12) <- "Isn't working with"
string

str_sub(string, -11) <- "really interesting."
string

# str_length("in reports!")

```

# Working with strings in datasets

```{r}
spotify_songs <- read_csv("data/spotify_sample.csv")
```

```{r}
spotify_songs %>% 
  mutate(track_name_lower = str_to_lower(track_name)) %>% 
  select(track_name, track_name_lower)
```

# 15 Min Task modifying spotify_songs

Upcase the text in the playlist_name column. Overwrite the existing column to store the result.
```{r}
spotify_songs <- spotify_songs %>% 
  mutate(playlist_name = str_to_upper(playlist_name))
spotify_songs
```

Lowcase the text in the track_artist column. Store the result in a new column called track_artist_lowcase.
```{r}
spotify_songs <- spotify_songs %>% 
  mutate(
    track_artist_lowcase = str_to_lower(track_artist)
  )
```

Take the first 5 characters of the playlist_id column, and save it in a new column called playlist_start.
```{r}
spotify_songs <- spotify_songs %>% 
  mutate(playlist_start = str_sub(playlist_id, 1, 5))
```

Take the first 4 characters of the track_album_release_date column and save it in a new column called year.
```{r}
spotify_songs <- spotify_songs %>% 
  mutate(year = str_sub(track_album_release_date, 1, 4))
```

Format the column danceability column to 2 significant digits (hint, look up ?format for how to use this). Overwrite the column with these values.
```{r}
spotify_songs <- spotify_songs %>% 
  mutate(danceability = format(danceability, digits = 2))
```

What else can you do to this dataset to practice? Can you work with the numbers? This is open ended - have fun!

```{r}
# No thanks, I'd rather eat lunch!
```

