---
title: "R Notebook"
output: html_notebook
---

# Missing Values

What we do with missing values depends on the context we're working in.

## Investigate the data to find out what type of missing value you have.

## Categories of missing values

### Missing Completely at Random MCAR
No pattern. We need to investigate if there is a pattern to the missing values

## Missing at Random
the probability of a value missing can be related to another variable. 

## Missing not at Random (MNAR)
When the missing values are dependent on the column, for example if people with lower incomes have not provided their income.

```{r}
is.na(NA)
```
```{r}
is.na(NaN)
```

# What to do with missing values

## Ignore?

## Drop?
Drop depending on category.
If you drop a MNAR you can create bias

## Replace?
Imputation - many options. 
some basic replacements are mean, median, mode.

# Read in data

```{r}
library(tidyverse)
```

```{r}
comms_data <- read_csv("data/telecom_data.csv")
```

# Identifying Missing Values

In one column

```{r}
comms_data %>% 
  filter(is.na(monthly_charges))
```

Check everywhere, cheat.
Omit drops all missing values from all places. We can then check the size of this against the original dimensions.

```{r}
comms_data %>% 
  na.omit() %>%
  dim()
```
```{r}
dim(comms_data)
```

Count missing values

```{r}
comms_data %>% 
  summarise(sum(is.na(monthly_charges)))
```

Remove missing values

```{r}
comms_data_no_missing <- 
  comms_data %>% 
  filter(!is.na(monthly_charges))
```

Replace missing Values - Impute

```{r}
comms_data %>% 
  mutate(monthly_charges = coalesce(monthly_charges, mean(monthly_charges, na.rm = TRUE)))
```
These methods won't pick up anything except NA and NaN values.

## Hidden Missing Values
But sometimes missing values have a value, represented by strings like "na" "N/A" "NN" "-99" "-999" "" (a blank string)
```{r}
comms_data %>% 
  filter(is.na(total_charges))
```

For categorical data

```{r}
comms_data %>% 
  mutate(total_charges = na_if(total_charges, "na"),
         total_charges = na_if(total_charges, "N/A"))
```

For numeric data
```{r}
comms_data %>% 
  mutate(total_charges = as.numeric(total_charges))
```

# Practice
Take a look at the payment_method column and find the different type of missing value identifiers we have, and how many we have.
Replace these with NA, and then check if you have the right amount

Find all values

```{r}
comms_data %>% 
  summarise(unique(payment_method))
```

Mutate values to NA using na_if

```{r}
comms_data %>% 
  mutate(payment_method = na_if(payment_method, "--")) %>% 
  summarise(unique(payment_method))
```

