---
title: "3_dplyr"
output: html_notebook
---

# Viewing

```{r}
library(tidyverse)
```

```{r}
all_deaths <- read_csv("data/character-deaths.csv")
```

```{r}
head(all_deaths)
tail(all_deaths)
```

```{r}
dim(all_deaths)
```

```{r}
names(all_deaths)
```

```{r}
glimpse(all_deaths)
```

```{r}
all_deaths
```

# Wrangling with dplyr

## Select : choose columns(variables) to keep

```{r}
# Selecting 3 columns, don't save
select(all_deaths, name, allegiances, year_of_death)

# Assigning the 3 columns to a new variable
three_column_death <- select(all_deaths, name, allegiances, year_of_death)

## Select all but name, don't save
select(all_deaths, -name)
```

# Practice
Select all columns which relate to a book

```{r}
# find column names
names(all_deaths)

# Deselect all without direct reference to book
select(all_deaths, -name, -gender, -allegiances, -nobility, -year_of_death)

#Using starts with

select(all_deaths, starts_with("book"))

#Using contains

select(all_deaths, contains("book"))

```

# Filter : Picking rows (observations) from data

```{r}

head(all_deaths, 3)
```


```{r}


## Only Lannister allegiance
filter(all_deaths, allegiances == "Lannister")

## All but Lannister allegiance
filter(all_deaths, allegiances != "Lannister")

```

## Check for value changes

```{r}
unique(all_deaths$allegiances)
```

Turns out we have allegiances assigned to Lannister and House Lannister

```{r}
## Anything which contains Lannister
filter(all_deaths, allegiances %in% c("House Lannister", "Lannister"))
```

## Numeric
```{r}

filter(all_deaths, year_of_death >= 299)
```

```{r}
filter(all_deaths, allegiances %in% c("House Lannister", "Lannister") & nobility == 1)

unique((filter(all_deaths, allegiances %in% c("House Lannister", "Lannister") & nobility == 1))$nobility)
```

## Practice

```{r}
filter(all_deaths, year_of_death <=299)

```

# Arrange : orders rows

Defaults to ascending

```{r}
# arrange wtih females first

arrange(all_deaths, gender)

# Descending

arrange(all_deaths, desc(gender))

# Ascending gender then descending nobility

arrange(all_deaths, gender, desc(nobility))
```

Find the females (gender is 0) who are not Lannisters

```{r}
filter(all_deaths, gender == 0 & !allegiances == c("Lannisters", "House Lannister" ))
```

Find just the data for the characters “Jon Snow”, “Daenerys Targaryen” and “Samwell Tarly”.


Arrange all_deaths by allegiances. What happens when you arrange by a character column?

```{r}
arrange(all_deaths, allegiances)
```

Arrange all_deaths by allegiances and book_intro_chapter
```{r}
arrange(all_deaths, allegiances, book_intro_chapter)
```

Arrange all_deaths by descending year_of_death
```{r}
arrange(all_deaths, desc(year_of_death))
```

# Mutate Function

## New Column

Create a column which contains years survived = year of death - 298
```{r}
mutate(all_deaths, years_survived = year_of_death - 298)
```

## Overwrite Existing Column

```{r}
mutate(all_deaths, book_of_death = book_of_death * 5)
```

# Summarise

calculates summary statistics

```{r}
summarise(all_deaths, females = sum(gender))
```

## Grouping

```{r}
# Groups all the zeros together and all the ones together
# More transparent than the above
sex_grouped <-  group_by(all_deaths, gender)

summarise(sex_grouped, count = n() )
```

```{r}
deaths_grouped <- group_by(all_deaths, allegiances, gender)

summarise(deaths_grouped, count = n())
```

# Practice

If a character doesn’t have a missing value for book_of_death, this means that they have died at some point. Use this to find the allegiances with the highest number of people dead.
Hint You will need to:
Filter the data so we just have characters who have died.

```{r}
have_died <- filter(all_deaths, !is.na(book_of_death)) 
```

Use group by and summarise to find the number of characters who belonged to each allegiance.

```{r}
dead_allegiance <- group_by(have_died, allegiances)

summarise(dead_allegiance, count=n())

dead_allegiance
```

# Pipes
and then. Start with data.

```{r}
have_died <- filter(all_deaths, !is.na(book_of_death)) 
died_grouped <- group_by(have_died, allegiances)
died_counts <- summarise(died_grouped, count = n())
```
 
```{r}
all_deaths %>% 
  filter(!is.na(book_of_death)) %>% 
  group_by(allegiances) %>% 
  group_by(allegiances) %>% 
  summarise(count = n()) %>% 
  arrange (desc(count))
```
 
 # Pull function
 
 makes something a value
 
```{r}
average_year_of_death <- all_deaths %>% 
  summarise (average_value = mean(year_of_death, na.rm = TRUE))

average_year_of_death # print to check
class(average_year_of_death) #data.frame

```
 
```{r}
average_year_of_death <- all_deaths %>% 
  summarise (average_value = mean(year_of_death, na.rm = TRUE)) %>% 
  pull()
average_year_of_death
class(average_year_of_death) # numeric
```
 
