---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
new_coders <- read_csv("data/new_coders.csv")
```

```{r}
new_coders %>% 
  group_by(reasons_to_code) %>% 
    mutate(
      mean_expected_earn = mean(expected_earn),
      difference_from_avg = expected_earn - mean_expected_earn
      )
```

Add a new column average_age_for_gender, which shows the average age of people the same gender as the respondent.

```{r}
new_coders %>% 
  select(gender, age) %>% 
  group_by(gender) %>% 
  mutate(average_age_for_gender = mean(age)
         )
```
# Stealing Damien's variance code!

```{r}
new_coders %>% 
  group_by(gender) %>% 
  mutate(
    average_age_for_gender = mean(age),
    variance = age - average_age_for_gender
  ) %>% 
  select(gender, age, average_age_for_gender, variance)
```

```{r}
library(CodeClanData)

chinesemeal
```

# Window Functions

## Lag and Lead

Lags and leads depend on ordering and grouping data

```{r}
chinesemeal %>% 
  arrange(Year) %>% 
  group_by(FoodType) %>% 
  mutate(
    last_year_calories = lag(CaloriesPerDay),
    difference_from_last_year = CaloriesPerDay - last_year_calories,
    five_years_ago_calories = lag(CaloriesPerDay, n=5) 
         ) %>% 
  select(Year, CaloriesPerDay, last_year_calories, five_years_ago_calories)
```

# Cumulative sum and mean
A cumulative mean can smooth your data out.

Find the cumulative average calories for each type of food in the chinesemeal data.

```{r}
chinesemeal %>% 
  group_by(FoodType) %>% 
  mutate(
    cumulative_calories = cummean(CaloriesPerDay)
  )
```

# Across - Doing stuff on multiple columns

```{r}
deaths = read_csv("data/character-deaths.csv")
```

```{r}
deaths %>% 
  mutate(across(book1_GoT:book5_DwD, as.logical))
```

```{r}
deaths %>% 
  summarise(across(book1_GoT:book5_DwD, sum))
```


# Where
```{r}
deaths %>% 
  summarise(across(where(is.numeric), sum))
```
# Logical Tests

is.character
is.logical etc.

# Across Practice

Change all character columns in new_coders into factors. Use across() to do this.

```{r}
deaths %>% 
  mutate(across(where(is.character), as.factor))
```

