---
title: "R Notebook"
output: html_notebook
---

# Lists

Acts like a container.

```{r}
x <- list(1, "a", TRUE, c(1,2,3))
```

```{r}
c(1:5)

as.list(1:5)

list(1:5)
```

```{r}
x[3]
# returns a list of one element
```


```{r}
x[[3]]
# a logical vector
```

# Named list

```{r}
named_list <- list(a = "I love R", b = 1:10)
```


```{r}
#These are all the same calls
named_list[[2]]
named_list[["b"]]
named_list$b
```

```{r}
names(named_list)
```

# Nested lists

```{r}
student_info <- list(
                "Agnes" = list(height = 1.8, shoe_size = 8L),
                "Bobby" = list(height = 2.1, gender = "Male"),
                "Charlie" = list(gender = "Female", likes_beets = TRUE)
)

student_info
```

Find Agnes height
```{r}
names(student_info)

names(student_info$Agnes)

student_info$Agnes$height
```

Find Agnes height assuming the list is unnamed.
```{r}
#Pirnt student info, Agnes is index 1
student_info

# Print Agnes, height is index 1
student_info[[1]]

student_info[[1]][[1]]
```

# Practice
Make a named list that contains a numeric vector and a character vector.

Access the second element of your list using double square brackets [[]]. What is it’s class?

Access the second element of your list using single square brackets []. What is it’s class?

Access the second element of your list by name. What is it’s class?
```{r}
practice <- list(num = 1:10, char = c("a character vector", "or two", "never three!"))
```

```{r}
practice[[2]]
class(practice[[2]])
# character
```

```{r}
practice[2]
class(practice[2])
# a list
```

```{r}
practice[["char"]]

class(practice[["char"]])
```

# Matrices

Matrices take values and build a table with them. Recycling values to fill the table.
It defaults to assigning values column by column.
```{r}
matrix(1:4, nrow =2, ncol =2)
```

```{r}
m <-matrix(1:4, nrow = 2, ncol = 2, byrow = TRUE)
```

# Selecting elements in a matrix

```{r}
m <- matrix(1:6, nrow = 2, ncol = 3)
```


We select objects in a matrix by row and column number, first row then column

Select everything in first row
```{r}
m[1, ]
```

Select everything in first column
```{r}
m[ ,1]
```

Select row 1, column 3
```{r}
m[1,3]
```

Select multiple rows
```{r}
m[ , c(1,3)]
```

# Calculations in matrices

We can manipulate everything in a matrix at once. apparently you can do this with vectors too!
```{r}
m*3
m-2
```

# Calculations on rows or columns

sum of rows returned as a numeric vector
```{r}
class(rowSums(m))
```

mean of columns returned as numeric vector
```{r}
colMeans(m)
```

colSums
rowMeans

# Practice
Make a matrix with 3 rows and 4 columns. Each column should contain the numbers 1 to 3.
```{r}
prac <- matrix(1:3, nrow = 3, ncol = 4)
prac <- matrix(rep(1:3), nrow = 3, ncol = 4)
```

Use dim() to check that the dimensions are as you expect.
```{r}
dim(prac)
```

Check that all the columns are identical by checking that their sums are the same.
```{r}
colSums(prac)
```


Name the rows of the matrix ‘one’, ‘two’ and ‘three’.

```{r}
rownames(prac) <- c("one", "two", "three")
```

