---
title: "R Notebook"
output: html_notebook
---

1 Introduction
Let’s practice drawing and analysing sampling distributions from a larger dataset that we will treat as a population. Slightly morbidly, we’re going to use data derived from a study of heart disease!

Remember throughout the following tasks that this is something of a ‘fake’ situation:
Normally, we would have a single sample of a given size drawn from a population, ideally containing as much data as possible.
Here, we’re treating a sample as if it were a population, and repeatedly drawing smaller samples from it. We’re doing this for educational purposes, to let you see what would happen if you could sample a population multiple times.


2 MVP
Load and explore the data held in heart.csv. A data dictionary is also available.

```{r}
library(tidyverse)
library(infer)

heart <- read_csv("data/heart.csv")
```

Plot a histogram and/or boxplot for thalach (maximum heart rate achieved in stress testing). How normal is the distribution?

```{r}
heart %>% 
  ggplot() +
  aes(x = thalach) +
  geom_boxplot()
```
1 outlier, median just over 150, left skewed

```{r}
heart %>% 
  ggplot() +
  aes(x = thalach) +
  geom_histogram(col = "white", bins = 25)
```
Left skewed, near normal dist, steeper drop on high thalach than low, poss outliers at ~40, unimodal, mode ~165

Also plot a barplot of exang (whether the patient displayed exercise induced angina)

```{r}
heart %>% 
  ggplot(aes(x = exang)) +
  geom_bar()
```
exang of 0 roughly twice as likely as exang of 1. Assume 1 = experienced exang and 0 = no exang reported

Now calculate population parameters: what is the mean thalach in the population. What proportion of patients have exercise induced angina (exang)?

```{r}
pop_param <- heart %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang == 1))
pop_param
```


Now take a single sample of 50-patients from the population and calculate point estimates for mean thalach and proportion of exang from this sample (let’s call these mean_thalach and prop_exang going forward).
[Hint - use rep_sample_n() from the infer package to do this]

```{r}
sample_50_1 <- heart %>% 
  rep_sample_n(size = 50, rep = 1, replace = F)
```

```{r}
pe_50_1 <- sample_50_1 %>% 
  ungroup() %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang == 1))

pe_50_1
```


What are the sampling errors of these point estimates?
[Hint - remember a sampling error is defined as the difference between a point estimate drawn from a sample, and a population parameter]

```{r}
pe_50_1 - pop_param
```

Now we’re going to repeatedly sample 50-patients from the population and build up sampling distributions for the mean_thalach and prop_exang point estimates. Take 5000 samples and store the sampling distributions in a variable rep_sample_50.
[Hint - again, use rep_sample_n() to do this.]

```{r}
sample_50_5000 <- heart %>% 
  rep_sample_n(size = 50, reps = 5000, replace = F)

pe_50_5000 <- sample_50_5000 %>% 
  ungroup() %>% 
  group_by(replicate) %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang == 1))

```

Visualise the sampling distributions of mean_thalach and prop_exang as histograms. How normal do the sampling distributions look?

# Mean thalach sampling distribution 
```{r}
pe_50_5000 %>% 
  ggplot() +
  aes(x = mean_thalach) +
  geom_histogram(col = "white")
```

# Prop Exang Sampling Distribution
```{r}
pe_50_5000 %>% 
  ggplot() +
  aes(x = prop_exang) +
  geom_histogram(col = "white", alpha = 0.8)
```

Obtain standard errors of the sampling distribution of mean_thalach and prop_exang directly as the standard deviations of the sampling distributions held in rep_sample_50

# Standard Error
Standard error is the sd of the sampling distribution

```{r}
se_50_5000 <- pe_50_5000 %>% 
  summarise(se_thalach = sd(mean_thalach),
            se_exang = sd(prop_exang))

se_50_5000
```


Repeat steps 7. and 9. (omitting step 8.) once more with samples of size 100-patients, and again with samples of size 200-patients. How do the standard errors of mean_thalach and prop_exang change as sample size increases? Does the uncertainty in the calculated statistics increase or decrease as sample size increases?
[Hint - it’s OK to copy, paste and alter code to do this!]

```{r}
sample_100_5000 <- heart %>% 
  rep_sample_n(size = 100, reps = 5000, replace = F)

pe_100_5000 <- sample_100_5000 %>% 
  ungroup() %>% 
  group_by(replicate) %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang == 1))

```

# Standard Error
```{r}
se_100_5000 <- pe_100_5000 %>% 
  summarise(se_thalach = sd(mean_thalach),
            se_exang = sd(prop_exang))
```


```{r}
sample_200_5000 <- heart %>% 
  rep_sample_n(size = 200, reps = 5000, replace = F)

pe_200_5000 <- sample_200_5000 %>% 
  ungroup() %>% 
  group_by(replicate) %>% 
  summarise(mean_thalach = mean(thalach),
            prop_exang = mean(exang == 1))

```

# Standard Error
```{r}
se_200_5000 <- pe_200_5000 %>% 
  summarise(se_thalach = sd(mean_thalach),
            se_exang = sd(prop_exang))
```


```{r}
se_50_5000
se_100_5000
se_200_5000
```
Standard error decreases as sample size increases.
Or certainty increases as sample size increases.