```{r}
library(sf)
library(tidyverse)
# library(ggplot2)
```
# Reading in spatial data
```{r}
north_carolina <- st_read(system.file("shape/nc.shp", package = "sf"))

class(north_carolina)
```

See the geometries
```{r}
nc_geometry <- st_geometry(north_carolina)
nc_geometry[[1]]
```

# Plot data
```{r}
plot(north_carolina)

# Plot one variable of one observation
plot(north_carolina[1, 1])

# Plot one variable of all observations
plot(north_carolina[ ,1])
```

# Task
Have a look through some of the variables within your north_carolina dataset, and see if you can create a spatial plot using the techniques above.
```{r}
plot(north_carolina["BIR74"])
plot(north_carolina["BIR79"])
plot(north_carolina[2])
plot(north_carolina[1:10, 9])
```

# Plotting sf in ggplot
```{r}
ggplot(north_carolina) +
  aes(fill = SID74) +
  geom_sf() +
  theme_bw()
```

```{r}
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
```

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```

# Just plot shape
```{r}
ggplot(world) +
  geom_sf(colour = "black", fill = "forest green")
```
# Plot with a variable
```{r}
ggplot(world) +
  aes(fill = pop_est) +
  geom_sf() +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

# Task
Recap your knowledge from ggplot week, and set your geom_sf aesthetic to be filled with the estimated gdp (gdp_md_est variable). Extra points if you make your map colour blind friendly!
What does your plot tell you? What does it tell you compared to the population?
```{r}
# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

world %>% 
  ggplot() +
  aes(fill = gdp_md_est) +
  geom_sf() +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt")
```

```{r}
italy <- world %>% 
  filter(name == "Italy")

ggplot(italy) +
  geom_sf() +
  ggtitle(italy$name)
```

# Calculaing distance

UTM - a system for assigning coordinates to a location on the earth
```{r}
# transform points from degrees to meters.
country_transformed <- st_transform(italy, 3003)
```


```{r}
# grid of region of interest
grid <- st_make_grid(country_transformed,
                     cellsize = 50000,
                     what = "centers")
plot(grid)
```


```{r}
# convert our shape from a polygon to points
country_line <- st_cast(country_transformed, "MULTILINESTRING")
```

```{r}
distance <- st_distance(country_line, grid)

distance
```

# Zoom in our map
```{r}
ggplot(world) +
  geom_sf(aes(fill = pop_est)) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE)
```

# Task
Have a play around with your coordinate systems, and try to plot a different region of the world. *Hint: read the help documentation for coord_sf.
```{r}
ggplot(world) +
  aes() +
  geom_sf(fill = "dark red") +
  coord_sf(
    xlim = c(110, 122),
    ylim = c(-37, -16.5),
    expand = FALSE
  ) +
  ggtitle("The West Coast of Australia")
```

# Labelling
```{r}
# Find centres of the shape (this gives a warning about calculating centres of longitude and latitude)
world_centres <- st_centroid(world$geometry)

# get the coordinates of these centres
world_coordinates <- st_coordinates(world_centres)

# add these to our data
world_points <- cbind(world, world_coordinates)
```

```{r}
ggplot(world_points) +
  geom_sf(fill = "grey") +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  geom_text(aes(x = X, y = Y, label = name), color = "darkblue", fontface = "bold")
```

