---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Making dataframes - customer details and orders

```{r}
customers <- tibble(
  customer_ID = c(001, 002, 004, 005, 008, 010), 
  name = c("John Smith", "Jane Adams", "Robert Landry", "Jane Tow", "Raul Tann", "Hilary Joyal"),
  email_address = c("johnsmith@gmail.com", "janea@gmail.com", "rlabdry@hotmail.com", "janet89@aol.com", "paul.tann1@gmail.com", NA),
  shipping_address = c("32 Station Road, Edinburgh", "42 Park Drive, Edinburgh", NA, "10 Gardiner Rd, Edinburgh", "12 Main St, Edinburgh", " 234 Queensferry Rd, Edinburgh,")
)
```

```{r}
orders <- tibble(
  customer_ID = c(001, 002, 005, 010, 012, 001), 
  order_number = c("KU65781", "SU98201", "JS76821", "YR28212", "JS91221","YR21873"),
  order_total = c("£32.50", "£56.00", "£100.02", "£20.25", "£39.50", "£45.00"),
  product_code = c("XZY780", "XYZ123", "XYZ256", "XYZ262", "XYZ345", "XYZ127")
)
```

# Combining these two dataframes
All the columns from one table, followed by all the columns from the second column, with only one repeat of the by = variable.
Different types of join give us different numbers of rows.

## Inner Join
Minimum number of rows. Returns the rows where the by = variable appears in both tables.

```{r}
inner_join(
  customers, orders,
  by = "customer_ID"
) %>%  view
```

## Left Join
Returns all of the rows from the left table (first table in the argument) and adds any matching information from the right hand table or NAs. If the right hand table has repeats, the output will have repeat rows for each repeat in the right hand table.

```{r}
left_join(
  customers, orders,
  by = "customer_ID"
) %>%  view
```

## RightJoin
as above with the argument positioned switched

```{r}
right_join(
  customers, orders,
  by = "customer_ID"
) %>%  view
```

## Full Join
The maximum number of joins, a left and a right join. Every row from the left hand table (at least once) and at least one row for every row in the right hand table.
```{r}
full_join(
  customers, orders,
  by = "customer_ID"
) %>%  view
```

## If primary and foreign key have different names

```{r}
inner_join(customers, orders, by = c("name_in_left" = "name_in_right"))
```

# Adding more tables so we have more overlapping columns

```{r}
customers2 <- customers %>% 
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "card"))
orders2 <- orders %>%
  mutate(payment_type = c("card", "cash", "cash", "card", "card", "cash"))
```

## Inner join by two things
Returns rows where two variables match, e.g. 
Useful when we don't have a useful primary key as we can make our primary key a set of two or more matching variables.

```{r}
inner_join(customers2, orders2, by = c("customer_ID", "payment_type"))
```
# Filtering joins
Things which aren't quite joins - less frequently used.
Filtering things in the first dataset depending on what's in the second.

## Semi Join
Only returns rows where the left hand table has a match in right hand, with no repeats (first instance selected).
In this instance, returns one row for every customer who has made an order.

```{r}
semi_join(
  customers, orders,
  by = "customer_ID"
) %>%  view
```

## Anti join
Returns any rows from first table, which don't appear in right hand table.
In this instance, we see the customers who have not placed an order

```{r}
anti_join(customers, orders, by = "customer_ID")
```

# 

```{r}
statuses <- tibble(
  status = c("active", "active", "inactive", "inactive", "active", "inactive")
) 

past_customers <- tibble(
  customer_ID = c(020, 021), 
  name = c("Dave Brown", "Jess Bishop"),
  email_address = c("dbrown@hotmail.com", "jessb@gmail.com"),
  shipping_address = c("57 Park View, Leith", "12 Reid Way, Broxburn")
)
```

## Bind cols
Adds all columns in right hand table to end of left hand table by position
```{r}
bind_cols(customers, statuses)
```

## Bind rows
Add rows from right hand table to the end of left hand table (only works when col_names() are the same)

```{r}
bind_rows(customers, past_customers)
```

