---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
actors <- read_csv("data/actors.csv")

movies <- read_csv("data/movies.csv")

roles <- read_csv("data/roles.csv")
```

1
List all the roles, along with the actors who played each role in The Little Mermaid. (In the data as “Little Mermaid, The”)

```{r}
# Link roles and movies

roles_movies <- inner_join(movies, roles, by = c("id" = "movie_id")) %>%  view

# link actors to roles and movies

roles_movies_actors <- inner_join(roles_movies, actors, by = c("actor_id" = "id")) %>% view

# select Little Mermaid data
roles_movies_actors %>% 
  filter(id == 192017) %>% 
  select(role, first_name, last_name)
```


2
Find the actor who has appeared in most films. List the films they appeared in, the year the film was released and their role in the film.

```{r}
# Find actor - actor is 22591	
roles_movies_actors %>% 
  group_by(actor_id) %>%
  summarise(num_film = length(unique(name))) %>% 
  arrange(desc(num_film))

# Films appeared in, year film release, role in film
roles_movies_actors %>% 
  filter(actor_id == 22591) %>% 
  select(name, year, role)
```


3
For all actors who have appeared in at least two films, 
find the average rating of the films they appeared. 

Find the five actors with the highest average rank of films they appeared in.

```{r}
# For all actors who have appeared in at least two films,
roles_movies_actors %>% 
  filter(film_count >= 2) %>% 
# find the average rating of all films they appeared in.
  group_by(actor_id) %>% 
  mutate(average_film_rating = mean(as.double(rank))) %>% 

# Find the five actors with the highest average rank of films they appeared in.
  arrange(desc(average_film_rating)) %>% 
  head(5)
```

