---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(prob)
library(tidyverse)
```

# Uniform Probability distribution
Where every outcome has the same likelihood of occuring. e.g. where our coins are not biased.

```{r}
s_three_coins <- tosscoin(3)
s_three_coins
```

```{r}
s_three_coins <- s_three_coins %>%
  mutate(label = str_c(toss1, toss2, toss3, sep = "")) %>%
  mutate(prob = 1 / n())

s_three_coins %>%
  ggplot(aes(x = label, y = prob)) +
  geom_col() + 
  xlab("coin toss result") + 
  ylab ("prob of each outcome")
```

# Plot number of heads
```{r}
s_three_coins <- s_three_coins %>% 
  mutate(
    num_heads = rowSums(. == "H")
  )

prob_num_heads <- s_three_coins %>% 
  group_by(num_heads) %>% 
  summarise(prob = sum(prob))

prob_num_heads %>% 
  ggplot() +
  aes(x = num_heads, y = prob) +
  geom_col() +
  xlab("Number of heads") +
  ylab("Probability")
```

# Measures of Centre

```{r}
library(janitor)
library(lubridate)
```

```{r}
air_con_sales <- read_csv("data/AirConSales.csv") %>% 
  clean_names() %>% 
  mutate(date = as_date(date, format = "%m/%d/%Y"))
```

# Find the distribution of units sold per day

```{r}
# col 2 = frequency e.g. 9 days where 3 units were sold
# Col 3 equals relative frequency: the proportion of days where 3 units were sold
sales_frequency_table <- air_con_sales %>% 
  tabyl(units_sold)

```

# Plot distribution of air con unit sales
```{r}
sales_frequency_table %>% 
  ggplot() +
  aes(x = units_sold, y = percent) +
  geom_col()
```

# Summarising distribution

## Mean

```{r}
air_con_sales %>% 
  summarise(mean_daily_sales = mean(units_sold))
```

## Median
The value where half of values are smaller than and half of the values are bigger than it
```{r}
air_con_sales %>% 
  summarise(median_daily_sales = median(units_sold))
```

## Mode
The value which occurs most frequently
```{r}
get_mode <- function(data){
  tabled_data <- table(data)
  table_names <- names(tabled_data)
  return( table_names[tabled_data == max(tabled_data)] )
}
```

```{r}
air_con_sales %>% 
  summarise(mode_daily_sales = get_mode(units_sold))
```

# Effects of outliers

```{r}
air_con_sales %>% 
  summarise(
    mean_daily_sales = mean(units_sold),
    median_daily_sales = median(units_sold),
    mode_daily_sales = get_mode(units_sold)
  )
```

Outliers have significant impact on mean value
```{r}
# Remove outlier
air_con_sales %>% 
  filter(units_sold < 43) %>% 
  summarise(
    mean_daily_sales = mean(units_sold),
    median_daily_sales = median(units_sold),
    mode_daily_sales = get_mode(units_sold)
  )
```

# Desciribing disributions

## Unimodal and Bimodal Distributions
Uniform - flat - no peaks
Unimodal - one most common value e.g. one peak in distribution
Bimodal - exactly two peaks
Multimodal - more than two peaks

```{r}
unimodal <- read_csv("data/unimodal.csv")
bimodal <- read_csv("data/bimodal.csv")
```

```{r}
unimodal %>% 
  ggplot() +
  aes(x = x) +
  geom_histogram(col = "white", bins = 20) 
```

```{r}
bimodal %>% 
  ggplot() +
  aes(x = x) +
  geom_histogram(col = "white", bins = 20)
```

# Measures of centre with bimodal or multimodal data
In the bimodal data set, the mean and median actually fall around 12, which is in fact a low point in the data.

```{r}
bimodal %>% 
  summarise(
    mean = mean(x),
    median = median(x)
  )
```

# Skewness
Skewed: The peak falls to the left or right of the centre
Skewness: a description of how skewed the data is, by looking at the relationship between mean and median
Left-skewed, the tail is on the left.

```{r}
left_skewed <- read_csv("data/leftskew.csv")
```

```{r}
left_skewed %>% 
  summarise(
    mean = mean(x),
    median = median(x)
  )
```

```{r}
left_skewed %>% 
  ggplot() +
  aes(x = x) +
  geom_histogram(col = "white", bins = 20)
```
# Calculating mode with continuous data
```{r}
left_skewed %>% 
  mutate(binned = cut(x, breaks = 20)) %>% 
  summarise(modal_value = get_mode(binned))
```

### Pattern for left skewed data
Mean < Median < Mode

# Right skewed
Try the same process again for the right-skewed data below:
right_skewed <- read_csv("data/rightskew.csv")
Visualise the distribution.
What order do the mean, median and mode fall in for right-skewed data?

```{r}
right_skewed <- read_csv("data/rightskew.csv")
```

```{r}
right_skewed %>% 
  ggplot() +
  aes(x = x) +
  geom_histogram(col = "white", bins = 20)
```

```{r}
right_skewed %>% 
  mutate(binned = cut(x, breaks = 20)) %>% 
  summarise(
    mean = mean(x),
    median = median(x),
    mode = get_mode(binned)
  )
```

Mean > Median > Mode

# Measuring skew

```{r}
library(e1071)
```

```{r}
left_skewed %>% 
  summarise(skewness = skewness(x, na.rm = T, type = 1))
```
Left skewness measured with a negative number
Right skewness measured with a positive number.
The greater the value of the number, the more skewed the data
0 > 0.5 - not skewed
0.5 > 1 - moderately skewed
1 > - highly skewed

```{r}
right_skewed %>% 
  summarise(skewness = skewness(x, na.rm = T, type =1))
```

```{r}
air_con_sales %>% 
  summarise(skewness = skewness(units_sold, na.rm = T, type = 1))
```

Impact of outliers on skewness
```{r}
air_con_sales %>% 
  filter(units_sold < 43) %>% 
  summarise(skewness = skewness(units_sold, na.rm = T, type = 1))
```

