---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
```

```{r}
jobs <- read_csv("data/TyrellCorpJobs.csv") %>% 
  clean_names()
```

# Range
the difference between maximum and minimum value

```{r}
# Range function calculates maximum and minimum values
jobs %>% 
  summarise(range_salary = range(salary))
```

```{r}
jobs %>% 
  summarise(
    range = max(salary) - min(salary)
  )
```

```{r}
jobs %>% 
  ggplot() +
  aes(x = salary) +
  geom_histogram(col = "white", bins = 25) +
  facet_wrap(~ position)
```

Outlier can greatly affect range
```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(range = max(salary) - min(salary))
```

# Interquartile Range

A method for trimming outliers

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(
    Q1 = quantile(salary, 0.25),
    Q2 = quantile(salary, 0.5),
    Q3 = quantile(salary, 0.75),
    IQR = Q3 - Q1
  )
```

# Five Number Summary

Min, Q1, median, Q3, Max of a distribution

```{r}
library(skimr)
```

```{r}
jobs %>% 
  select(-x1) %>% 
  group_by(position) %>% 
  skim()
```

# Box plot
A visualisation of a 5 number summary.
Whiskers are 1.5 times interquartile range. Anything outwith the whiskers is an outlier

```{r}
set.seed(42)
distribution_1 <- tibble(
  y = append(rnorm(n = 100, mean = 5, sd = 10), rnorm(n = 200, mean = 5, sd = 1))
)
distribution_2 <- tibble(
  y = runif(n = 1000, min = -30, max = 30)
)
```

Investigate and comment on the centrality and spreads of distribution_1 and distribution_2 produced by the following function calls.
Use geom_boxplot() and skim() first
Next, plot histograms to confirm your descriptions.
[Don’t worry about what the functions generating distribution_1 and distribution_2 are doing]
```{r}
distribution_1 %>% 
  skim()
```


```{r}
distribution_1 %>% 
  ggplot() +
  aes(y = y) +
  geom_boxplot()
```

Left skewed, quite a lot of outliers, relatively dense range in interquartile range.

```{r}
distribution_2 %>% 
  skim()
```

```{r}
distribution_2 %>% 
  ggplot() +
  aes(y = y) +
  geom_boxplot()
```

No outliers, not skewed, larger IQR, smaller range

```{r}
library(e1071)
heavily_right_skewed <- read_csv("data/heavily_right_skewed.csv")
```

```{r}
heavily_right_skewed %>% 
  summarise(skewness = skewness(x, type = 1))
```

```{r}
heavily_right_skewed %>%
  ggplot(aes(y = x)) +
  coord_flip() +
  geom_boxplot()
```

# Variance
The lower the number the lower the spread

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(variance = var(salary))
```

# Standard Deviation
Square root of variance.
The lower the number, the lower the spread

```{r}
jobs %>% 
  group_by(position) %>% 
  summarise(sd = sd(salary))
```

