---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Uniform Distributions

```{r}
dice <- tibble(
    x = 1:6,
    f_x = replicate(6, 1/6)
  ) %>%
  mutate(F_x = cumsum(f_x))
dice %>%
  ggplot(aes(x = x, y = F_x)) +
  geom_step() +
  xlab("number rolled on die") + ylab("probability") +
  scale_x_continuous(breaks = 1:6) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  ylim(0, 1) 
```

```{r}
brain_breaks <- tibble(
  l = seq(4, 23, by = 0.001),
  f_l = dunif(x = l, min = 5, max = 22)
)
brain_breaks %>%
  ggplot(aes(x = l, y = f_l)) +
  geom_line() +
  ylim(0, 0.075) + 
  xlab("l (minutes)") + ylab("f_l (probability density)") +
  geom_ribbon(aes(ymin = 0, ymax = ifelse(l >= 8.4 & l <= 10.751, f_l, 0)), fill = "red", alpha = 0.6)
```

```{r}
brain_breaks <- brain_breaks %>%
  mutate(F_l = punif(q = l, min = 5, max = 22))
brain_breaks %>%
  ggplot(aes(x = l, y = F_l)) +
  geom_line()
```

# Probability of a break between a range lengths

Probabilty of the longer one minus the probability of the shorter one

```{r}
punif(q = 10.751, min = 5, max = 22) - punif(q = 8.4, min = 5, max = 22)
```

# Normal Distributions

```{r}
three_norms <- tibble(
  x =  seq(0, 20, 0.1),
  f1_x = dnorm(x = x, mean = 10, sd = 1), 
  f2_x = dnorm(x = x, mean = 10, sd = 2),
  f3_x = dnorm(x = x, mean = 10, sd = 3)
)
three_norms %>%
  ggplot() +
  geom_line(aes(x = x, y = f2_x), col = "black") +
  geom_line(aes(x = x, y = f1_x), col = "red") +
  geom_line(aes(x = x, y = f3_x), col = "blue")
```

```{r}
library(janitor)
```

```{r}
jobs <-read_csv("data/TyrellCorpJobs.csv") %>% 
  clean_names()
```

```{r}
accounts_salary_stats <- jobs %>% 
  filter(position == "Accounting") %>% 
  summarise(
    num = n(),
    mean = mean(salary),
    sd = sd(salary)
  )

accounts_salary_stats
```

```{r}
jobs %>% 
  filter(position == "Accounting") %>% 
  ggplot() +
  aes(x = salary) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm,
    args = list(
      mean = accounts_salary_stats$mean,
      sd = accounts_salary_stats$sd
    ),
    col = "steel blue"
  )
```

# Z score
how many standard deviations is this value from the mean?
```{r}
management_scaled <- jobs %>% 
  filter(position == "Management") %>% 
  mutate(z_salary = scale(salary)) %>% 
  arrange(desc(z_salary))

management_scaled
```

```{r}
management_scaled %>% 
  filter(!between(z_salary, left = -3, right = 3))
```
z of more than 3 standard deviations positive or negative means something we want to explore and understand. Is it an error?

# Cumulative Distribution for normal distribution
The empirical 3 sigma rule.
The 68, 95, 99.7 rule

```{r}
standard_normal <- tibble(
  z = seq(from = -4, to = 4, by = 0.01),
  f_z = dnorm(x = z)
)
standard_normal %>%
  ggplot(aes(x = z, y = f_z)) +
  geom_line()
```

```{r}
pnorm(q = 0)
```

Chance of something lying within 1 standard deviation
```{r}
pnorm(q = 1) - pnorm(q = -1)
```

Chance of something lying within 2 standard deviation
```{r}
pnorm(q = 2) - pnorm(q = -2)
```

Chance of something lying within 3 standard deviation - 99.7%
```{r}
pnorm(q = 3) - pnorm(q = -3)
```
