---
title: "R Notebook"
output: html_notebook
---

```{r}
library(lubridate)
library(tidyverse)
library(nycflights13)

date1 <- ymd_hms("2017-04-22 22:01:18")
date2 <- ymd_hms("2001-11-07 10:09:56")
```

Question 1
Extract the different components (year, month, mday, yday, wday) from the variable date1. Add the labels to the month and do not abbreviate the month labels.

```{r}
year(date1)
month(date1, label = T, abbr = F)
mday(date1)
yday(date1)
wday(date1, week_start = 1)
```

Question 2
Add 14 days to your date1 variable and store it in a variable called next_fortnight. Take away 10 years from your date2 variable, and store it in previous_decade.
```{r}
next_fortnight <- date1 + days(14)
previous_decade <- date2 - years(10)
```

Question 3
Create a variable called this_year which contains today’s date. Then create an interval object which uses this_year and the previous_decade variable you made above. Store it in interval_years. Hint: you’ll need to use lubridate::interval(), as tsibble has an interval() function too
```{r}
this_year <- today()
interval_years <- interval(this_year, previous_decade)
```

Question 4
Change the time zone of both your date variables to “America/New_York”, and call them nyc_date1 and nyc_date2.
```{r}
nyc_date1 <- with_tz(date1, "America/New_York")
nyc_date2 <- with_tz(date2, "America/New_York")
```

Question 5
Use the flights data from the nycflights13 dataset. Make a date-time column called departure_date from the year, month, day, hour, and minute variables in the flights tibble.
```{r}
flights %>% 
  mutate(departure_date = make_datetime(
    year = year, month = month, day = day, hour = hour, min = minute
  ))
```