---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```


# Naming timezones
Olson names format: Continent/City

```{r}
Sys.timezone()
```

```{r}
OlsonNames()[1:10]
```

```{r}
length(OlsonNames())
```

tz extracts time zone
```{r}
tz(ymd_hms("2016-06-02 04:00:00"))
```

# Adding a timezone
```{r}
ymd_hms("2016-06-02 04:00:00", tz = "Pacific/Auckland")
```

```{r}
nz_date <- ymd_hms("2016-06-02 04:00:00", tz = "Pacific/Auckland")
```

# Converts timezone
```{r}
with_tz(nz_date, "Europe/London")
```

Task
Let’s say a plane leaves Sydney Australia at 2.15pm on the 7th August 2018, and arrives at 5.40am on the 8th August 2018 in London.
Create these two dates with the appropriate time zones.
Work out how long the duration between take off and landing is.
Say you left London at 6.20am on the 10th August 2018, and the total flight time including stop-overs was 28 hours and 46 minutes. At what time and date would you arrive in Sydney, Australia?

```{r}
departure <- ymd_hm("2018-Aug-7 14:15", tz = "Australia/Sydney")
arrival <- ymd_hm("2018-aug-8 5:40", tz = "Europe/London")
```

```{r}
as.period(arrival - departure)
```

```{r}
departure2 <- ymd_hm("2018-aug-10, 06:20", tz = "Europe/London")

arrival2 <- ymd_hms(departure2 + 
                    dhours(28) + 
                    dminutes(46), 
                    tz = "Australia/Sydney")

arrival2 <- with_tz(departure2 + hours(28) + minutes(46), "Australia/Sydney")
arrival2
```

