---
title: "R Notebook"
output: html_notebook
---

There is a baseR way of dealing with timeseries, dealing with objects of ts class.

There are a range of packages in the tidyverse which deal with time series in a tidy format

```{r}
library(tsibble)
library(tsibbledata)
library(tidyverse)
library(lubridate)
```


# Tsibble
tsibble is a tibble for time series data

Half hourly record of electricity demand in Victoria, Australia

```{r}
vic_elec
```

tsibbles can have an index, which is either timedate or date or time
find index variable
```{r}
index_var(vic_elec)
```

tsibbles can have a key, 
```{r}
global_economy
# measures for different countries throughout time. the key shows us that there are multiple time series in a dataframe, e.g. Afghanistan measurement is one time series in this database of multiple time series
key_vars(global_economy)
```

We can use tidyverse syntax on tsibble
```{r}
vic_elec %>% 
  select(Time, Demand, Temperature)

vic_elec %>% 
  # keeps time because it is the index
  select(Demand, Temperature)
```

# Additional functions in tsibbles

Filter index
```{r}
vic_elec %>% 
  filter_index("2012-01")

vic_elec %>% 
  filter_index("2012-01-01")

vic_elec %>% 
  filter_index("2012")
```

Add year column to tsibble
```{r}
vic_elec <- vic_elec %>% 
  mutate(year = lubridate::year(Time))
```

# Group by are more complicated because we have to preserve the index
```{r}
# Index by replaces the index with the new index variable
vic_elec %>% 
  index_by(year) %>% 
  summarise(avg_temp = mean(Temperature))

vic_elec %>% 
  index_by(Date) %>% 
  summarise(avg_temp = mean(Temperature))
```

# Plotting time series
Often needs a group and summarise first as too much data will build an unreadable plot.
```{r}
vic_elec %>% 
  index_by(Date) %>% 
  summarise(mean_temp = mean(Temperature)) %>% 
  ggplot()+
  aes(x = Date, y = mean_temp) +
  geom_line()
```

# Rolling window statistics
A different way of summarising/aggregating data
Window of 10: calculate for 1:10, 2:11, 3:12, 4:13, 5:14. 
Gives us a similar number of values to what we started with, but smoothes out fluctuations - helps us to see trend.
We also have cumulative mean.

A moving average using a window of 2000
```{r}
library(slider)

vic_elec <- vic_elec %>% 
  mutate(
    temp_moving_avg = slide_dbl(
      Temperature,
      # functional notation, with . standing for input
      ~ mean(., na.rm = TRUE),
      .before = 1000,
      .after = 1000
    )
  )
```

Plot moving average
```{r}
vic_elec %>% 
  ggplot() +
  aes(x = Time) +
  geom_line(aes(y = Temperature), col = "grey") +
  geom_line(aes(y = temp_moving_avg), col = "red")
```

Task - moving average with window size of 200 
More impacted by daily trends
```{r}
vic_elec <- vic_elec %>% 
  mutate(
    temp_moving_avg_200 = slide_dbl(
      Temperature,
      # functional notation, with . standing for input
      ~ mean(., na.rm = TRUE),
      .before = 100,
      .after = 100
    )
  )
vic_elec %>% 
  ggplot() +
  aes(x = Time) +
  geom_line(aes(y = Temperature), col = "grey") +
  geom_line(aes(y = temp_moving_avg_200), col = "red")
```

Task - add .complete = TRUE
Only calculate for the window if there are the full amount of datapoints available.
```{r}
vic_elec <- vic_elec %>% 
  mutate(
    temp_moving_avg_200 = slide_dbl(
      Temperature,
      # functional notation, with . standing for input
      ~ mean(., na.rm = TRUE),
      .before = 1000,
      .after = 1000,
      .complete = TRUE
    )
  )
vic_elec %>% 
  ggplot() +
  aes(x = Time) +
  geom_line(aes(y = Temperature), col = "grey") +
  geom_line(aes(y = temp_moving_avg_200), col = "red")
```

