---
title: "R Notebook"
output: html_notebook
---

# R has 6 classes for time and date

## BaseR
- Date - dates
- POSIXct - dates and times
- POSIXlt - dates and times, but allows extraction of indi components

## Tidyverse
- date - dates, allows extracting
- time - times, allows extracting
- dttm - dates and times, allows extracting

Lubridate lib (part of Tidyverse)
```{r}
library(lubridate)
```
Get today's date.
The way things are stored and the way things are presented are not always the same.
```{r}
today()
```
```{r}
class(today())
```

```{r}
now()
```
```{r}
class(now())
```

# Reading in and converting times and dates

## Strings
Most common problem
Lubridate has 3 functions depending on the order of the date in the string
 - ymd
 - mdy
 - dmy
 
Convert from class string to Date.

```{r}
ymd("2019-01-19")
mdy("January 9th, 2019")
class(dmy("09-Jan-2019"))

# For non-English language, change locale
dmy("09 janvier 2019", locale = "")
# accepts numeric
ymd(20190119)
```

```{r}
ymd_hms("2019-01-09 10:08:26")
class(ymd_hms("2019-01-09 10:08:26"))
```

```{r}
ymd_hm("2019-01-09 10:08")
ymd_h("2019-01-09 10")
```

Remember that _hm and _h attach 00 mins or seconds, which can sometimes cause problem.

Task
```{r}
mdy("March 12th, 2005")
ymd("2018-06-23")
ymd_hms("2017-04-29 11:43:02")
dmy("02-Sep-2002")
# "2001-01-07 7" I would set this to year 2001 unless I knew the format.  
mdy_hm("August 1st, 2019 08:13")
```

as_date = bad practice

## Components
```{r}
library(tibble)
library(tidyverse)

# create some date component data to practice with 
date_components <- tibble(id = 1:5, year = 2013 , month = 3:7, day = 1, hour = 5, minute = c(0, 5, 15, 43, 12))
date_components
```
```{r}
date_components %>% 
  mutate(
    date = make_datetime(
      year = year, 
      month = month, 
      day = day, 
      hour = hour, 
      min = minute)
  )
```

Task

```{r}
library(nycflights13)
data(flights)
```

Make a column called departure which makes a column of departure date and time
```{r}
flights %>% 
  mutate(departure = make_datetime(
    year = year,
    month = month,
    day = day,
    hour = hour,
    min = minute
  ))
```

## Existing date time objects

```{r}
class(as_datetime(today()))
```

```{r}
as_date(now())
```

Task
```{r}
last_event <- "2019-07-01 04:23:12"
```

Turn the variable last_event into a date-time, and then extract just the date component.
```{r}
# Make a date time from character vector
last_event <- ymd_hms(last_event)
class(last_event)

# Extract date from datetime
date(last_event)
as_date(last_event)
```

# Extracting date time components

```{r}
year(today())
# label = T -> return label not number, abbr = F -> return full name, not abbreviation
month(today(), label = T, abbr = F)
# Day of the month
mday(today())
# Day of the year (ordinal day)
yday(today())
wday(today(), week_start = 1, label = T, abbr = F)
date(now())
hour(now())
minute(now())
second(now())
```

Task
```{r}
our_datetime <- ymd_hms("2020-02-29 21:04:56")

year(our_datetime)
month(our_datetime, label = T)
mday(our_datetime)
yday(our_datetime)
wday(our_datetime, label = T, abbr = F)
date(our_datetime)
hour(our_datetime)
minute(our_datetime)
second(our_datetime)
```

# Time spans
4 classes of time span:
Period - e.g. years(1) = 1y 0m 0d 0H 0M 0S.
  Takes into account leap year and time difference
Duration - e.g. dyears(1) = 31557600s
Difftime
Interval

durations, which represent an exact number of seconds.
periods, which represent human units like weeks and months.
intervals, which represent a starting and ending point.

One year after today
```{r}
today() + years(1)
today() + dyears(1)
```

One year after 1st June 2020
```{r}
ymd("2020-01-01") + years(1)
ymd("2020-01-01") + dyears(1)
```

```{r}
now() + months(6)
now() + dmonths(6)

now() + days(150)
now() + ddays(150)
```

```{r}
# difftime
today() - ymd("1990-06-26")
class(today() - ymd("1990-06-26"))

# period
as.period(today() - ymd("1990-06-26"))
class(as.period(today() - ymd("1990-06-26")))

# duration
as.duration(today() - ymd("1990-06-26"))
class(as.duration(today() - ymd("1990-06-26")))
```

Task
Let’s use the date ‘25-12-2021’ (Christmas this year). Calculate the following:
```{r}
xmas_2021 <- dmy(251221)
```
How long is it from now until Christmas?
341 days
29462400s (~48.71 weeks)
```{r}
xmas_2021 - today()
```
```{r}
as.duration(xmas_2021 - today())
```

What date would 6 months after Christmas be?
2022-06-25
```{r}
xmas_2021 + months(6)
```
```{r}
xmas_2021 + dmonths(6)
```

What date would a fortnight before Christmas be?
2021-12-11
```{r}
xmas_2021 - weeks(2)
```
```{r}
xmas_2021 - dweeks(2)
```

What does adding on 36 hours after Christmas give you?
"2021-12-26 12:00:00 UTC"
```{r}
xmas_2021 + hours(36)
```
```{r}
xmas_2021 + dhours(36)
```

## Interval
the most precise type of time span, we give it a start and an end.

Make an interval
```{r}
job_start <- ymd("2019-06-21")
job_end <- ymd("2020-06-21")

job <- interval(job_start, job_end)
job_start %--% job_end
class(job)
```

manipulate intervals
```{r}
int_start(job)
int_end(job)
int_flip(job)
int_shift(job, by = ddays(1))
int_shift(job, by = weeks(10))
```

compare intervals
```{r}
biking_club <- interval(ymd(20190621), ymd(20200101))
university <- interval(ymd(20041001), ymd(20080601))
```

```{r}
# Does any one day overlap in these intervals?
int_overlaps(job, biking_club)
# Do any start or ends in one align with start or end point of the other?
int_aligns(job, biking_club)
```

Task
```{r}
datetime1 <- ymd_hms("2011-06-04 12:03:01") 
datetime2 <- ymd_hms("2011-07-25 14:24:06")

datetime3 <- ymd_hms("2011-07-20 13:23:21")
datetime4 <- ymd_hms("2011-08-31 15:02:11")

interval1 <- interval(datetime1, datetime2)
interval2 <- interval(datetime3, datetime4)
```

Use the setdiff() function to find the part of interval1 that is not in interval2.
Is interval1 %within% interval2? If not, make an interval3 that is %within% interval2

```{r}
# What of first argument is not contained in second argument?
setdiff(interval1, interval2)
setdiff(interval2, interval1)

# are all the dates of first argument contained within the second argument?
interval1 %within% interval2

interval3 <- interval(ymd("2011-07-30"), ymd("2011-08-15"))
interval3 %within% interval2
```

