---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(janitor)
library(modelr)
library(broom)
```

Load and clean the names of mortgage_applications.csv.
```{r}
mortgage <- read_csv("data/mortgage_applications.csv") %>% 
  clean_names()
```

Re-run the logistic regression on the dataset, again treating the tu_score, age and employed variables as predictors and the accepted variable as the binary dependent.

```{r}
mortgage_3pred_model <- glm(accepted ~ tu_score + age + employed,
             data = mortgage,
             family = binomial(link ="logit"))
```

Keep the model object (call it mortgage_3pred_model) and use it to predict estimated probabilities for the sample data using add_predictions() (call this mortage_data_with_3pred)
Look as the head() of the data with predicted probabilities.
```{r}
mortgage_data_with_3pred <- mortgage  %>% 
  add_predictions(mortgage_3pred_model, type = "response")

head(mortgage_data_with_3pred, n = 15)
```

```{r}
glance(mortgage_3pred_model)
```

# Threshold Probability

```{r}
# Set threshold
threshold <- 0.6

# Modify dataframe
mortgage_data_with_3pred <- mortgage_data_with_3pred %>% 
  mutate(pred_thresh_0.6 = pred >= threshold)

# View results
head(mortgage_data_with_3pred)

# We see that we will get errors
```

# Confusion Matrix
True Negatives, False Positives,
False Negatives, True Positives

```{r}
confusion_matrix <- mortgage_data_with_3pred %>% 
  tabyl(accepted, pred_thresh_0.6)

confusion_matrix
```
NTP <- 179
NTN <- 679
NFP <- 49
NFN <- 93

True Positive
```{r}
mortgage_data_with_3pred %>% 
  filter(tu_score == 631)
```
```{r}
mortgage_data_with_3pred %>% 
  filter(tu_score == 594)
```

# Accuracy = (nTP + nTN) / n

Accuracy is not a sufficient measure in and of itself.
Accuracy is a problem when our dataset is unbalanced because an incorrect threshold can get a high accuracy score.

# Rates
```{r}
NTP <- 179
NTN <- 679
NFP <- 49
NFN <- 93

TPR <- NTP / (NTP + NFN)
TNR <- NTN / (NTN + NFP)
FPR <- NFP / (NFP + NTN)
FNR <- NFN / (NFN + NTP)
```

