---
title: "R Notebook"
output: html_notebook
---

```{r, warning = FALSE}
library(tidyverse)
library(janitor)
library(mosaicData)
library(mosaic)
library(GGally)
library(ggfortify)
```

```{r}
head(RailTrail)
```

# Cleaning

```{r}
railtrail_clean <- RailTrail %>% 
  clean_names() %>% 
    mutate(
      spring = ifelse(spring == 1, TRUE, FALSE),
      summer = ifelse(summer == 1, TRUE, FALSE),
      fall = ifelse(fall == 1, TRUE, FALSE)
  )
```

# Redundant Predictors
Variables which can be predicted from, or are aliases of, other variables.
If we leave these in we will have multicolinearity and our model will account for these variables repeatedly.

```{r}
alias(lm(volume ~ ., data = railtrail_clean))
```

```{r}
railtrail_clean <- railtrail_clean %>% 
  select(-c("hightemp", "lowtemp", "fall", "day_type"))
```


# GGpairs
examine plots to find predictors which appear to be associated with volume. The outcome variable for our model.

```{r, warning = FALSE}
# library(GGally)
ggpairs(railtrail_clean)
```
This shows some correlation between volume and average temperature. So we can start building a model with the hypothesis that volume increases as avg_temp increases.

# Linear Regression Model
response variable: volume
predictor variable: avgtemp

volume = intercept + b(avg_temp) * avgtemp

## Plot lm model
```{r}
railtrail_clean %>% 
  ggplot(aes(x = avgtemp, y = volume)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Build simple linear regression model
```{r}
model <- lm(volume ~ avgtemp, data = railtrail_clean)
```

## Check Goodness of Fit
```{r}
# library(ggfortify)
autoplot(model)
```
```{r}
summary(model)
```
some deviation from normality in q-q, heteroscadacity in scale_location (this line slopes up rather than sitting horizontal to the x axis)
High p-value.
Residual standard error of 115.9 people!

# A parralel slopes model.
e.g. Add predictors to the model

## Find and check a variables significance to the output
We hypothesise that weekday has an effect on volume

```{r}
railtrail_clean %>% 
  ggplot(aes(x = weekday, y = volume)) +
  geom_boxplot()
```
Difference in numbers.
```{r}
railtrail_clean %>% 
  summarise(cor = cor(weekday, volume))
```
A weak negative correlation.

## Add to the model
# Adding a categorical variable
```{r}
model2 <- lm(volume ~avgtemp + weekday, data = railtrail_clean)
```

## Check effectiveness
```{r}
autoplot(model2)
```
```{r}
summary(model2)
```
r^2 has improved from 0.18 to 0.25.
WeekdayTRUE coefficient -70, means that every day which is a weekday we should see 70 less people visiting.
Pr(>|t|) = the p value for this test indicates that it is statistically significant.
avgtemp coefficient 4.5 means for each unit of temperature increase, we expect 4.5 more visitors.

```{r}
# library(mosaic)
plotModel(model2)
```
In a parallel slopes model, the only difference in calculating the line is the intercept. The lines are parallel.

# Interactions
Why do we need them?
```{r}
railtrail_clean %>%
  ggplot(aes(x = avgtemp, y = volume, color = weekday)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
avgtemp affects the volume differently depending on whether it is a weekday or not.

## Add interaction to the model
```{r}
# volume varies with average temperature and weekday and average temperature interacting with weekday
model4 <- lm(volume ~ avgtemp + weekday + avgtemp:weekday, data = railtrail_clean)

# another notation option. reads volume varies with avgtemp and weekday and all possible interactions between them
# model4 <- lm(volume ~ avgtemp *weekday, data = railtrail_clean)
```

## Plot new model
```{r}
plotModel(model4)
```
## Check whether the addition is justified.
```{r}
autoplot(model4)
```
```{r}
summary(model4)
```

# Adding a continuous variable

## Are cloudcover and volume associated?
```{r}
railtrail_clean %>% 
  ggplot(aes(x = cloudcover, y = volume)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
## Add to model
```{r}
model5 <- lm(volume ~ avgtemp + weekday + cloudcover, data = railtrail_clean)
autoplot(model5)
summary(model5)
```
R^2 increases to 0.4
Residual standard error drops by 11 to 100.3

## Plotting this model
Things get complicated. You could explore plotly package to draw 3D plots.

# Interaction with continuous predictors

We hypothesise that the relationship between avgtemp and precip will affect volume. e.g. a hot wet day will discourage users

# Check significance of interactions
Coplot, shingles plot or rooftiles plot
```{r}
# plot volume against avgtemp, conditional on precip
# volume versus avgtemp given precip
coplot(volume ~ avgtemp | precip, 
       # add line of best fit
       panel = function(x, y, ...) {
         points(x,y)
         abline(lm(y ~ x), col = "blue")
       },
       # set how much the precip shingles overlap
       overlap = 0.2,
       data = railtrail_clean)
```
This example is of a coplot where the interaction does not have a significant impact.
If the slope of the lines varied for each shingle, this would indicate a potentially significant interaction between the predictor on the x axis and the conditional predictor.

* the bottom left scatter plot corresponds to volume versus avgtemp for precip values in a narrow range from 0 inches to approximately 0.01 inches (i.e. essentially the days where it did not rain).
* the bottom right scatter plot corresponds to volume versus avgtemp for precip values in a broader range from 0 inches to approximately 0.16 inches.
* the top left scatter plot corresponds to volume versus avgtemp for precip values in a very broad range from approximately 0.14 inches to approximately 1.5 inches.
Why have the shingles ended up being the width that they are? R tries to arrange the shingles so that, as far as possible, they contain near equal numbers of data points.

```{r}
model7 <- lm(volume ~ avgtemp + cloudcover + weekday + precip + avgtemp:precip, data = railtrail_clean)

autoplot(model7)
```
```{r}
summary(model7)
```
This addition is not justified. without avgtemp:precip, precip gets a p value of 0.008, adding the avgtemp:precip is not statistically significant and it nullifies the effect of precip on the model.
