---
title: "R Notebook"
output: html_notebook
---

```{r}
library(stringr)
library(CodeClanData)
```

# If statements

Pseudo
```{r}
if (expression){
  do something
}

if (weather == "sunny") {
  hang_up_washing()
}
```

```{r}
if (10 < 15) {
  print("Yes")
}
```

```{r}
if (5 %in% c(1, 4, 8)){
  print("Yes")
}
```

```{r}
if (5 != 3){
print("Yes")
}
```

```{r}
x <- 8
y <- 7

if (x>y){
  print("Yes")
}
```

# Nested Ifs

```{r}
x <- 5
y <- 10

if (x %in% c(3, 4, 5)){
  
  print("x is in our list")
  
  if (x == 5){
    
    print("x is exactly 5")
    
  }
  
} else {
  
  print("Start of else statement")
  
  if (y %% x == 0){
    
    print("x divides y")
  
  } else(
    
    print("x does not divide Y")
  )
  
  print ("End of else statement")
```


# If Else

pseudo
```{r}
if (expression){
  do something when expression is true
} else {
  do something when expression is false
}

if (weather == "Rain"){
  wear_wellies()
} else {
  wear_trainers()
}
```

Build an if else statement so that it prints ‘Yes that divides’ when x is divisible by 2, and prints ‘No that does not divide’ otherwise.

```{r}
x <- 19

if (x %% 2 == 0){
  print("Yes that divides by 2")
} else {
  print("No that does not divide by 2")
}
```



# For Loops

Pseudo
```{r}
for ("item" in list/vector) {
  do somethething with "item"
}
```

```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for(day in weekdays) {
  print(str_c("The day is ", day))
}
```

```{r}
num <- c(1:5)

for(num in num) {
  print(num)
}
```
# Nested

```{r}
for (row_index in 1:5){
  for (col_index in 1:5){
    if (row_index == col_index){
      print(example_matrix[row_index, col_index])
    }
   }
}
```

# Practice
Create a nested for loop and if statement combination which loops through the weekdays and prints out the phrase ‘Last weekday!’ once it gets to Friday.


```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

for (day in weekdays) {
  if (day == "Friday") {
    print("Last weekday!")
  }
}
```


# Practice
Using the starwars list from the CodeClanData package, do the following:
Print the name for every character in the list
```{r}

for (item in starwars) {
  print(item$name)
}

```

Print out the hair_color of every female in the list.

```{r}
for (item in starwars) {
  if (item$gender == "female") {
    print(item$hair_color)
  }
}
```

(Harder) Print out all the URLs for the films of every character that is under 170cm tall.

```{r}
for (char in starwars) {
  if (as.double(char$height) > 170) {
    print(char$films)
  }
}
```


```{r}
str(starwars)
```

