---
title: "R Notebook"
output: html_notebook
---

# Tidy data Format

## 3 Rules

1. One variable = One column
2. One observation = One row
3. One value = One cell

Sounds easy, but it can be hard to define one variable or one observation.

In a survey, every question is one variable, every questionee is one observation

## Long format

## Wide format

# How to

```{r}
library(tidyverse)
```
```{r}
subject_hr <- tibble( name = c("SUBJ01","SUBJ02","SUBJ03"), 
  hr_am = c(68,72,68), 
  hr_pm = c(58,89,52))
```


Pivot longer takes data, columns you want to pivot, variable name for new columns and value to put in new column.

It moves items from the columns you specified, and takes the names of these columns into a new variable specified by variable name for new column and the values from these columns into a new variable whose nam is specified by values_to.
```{r}
subject_hr %>% 
  pivot_longer(
    col = c(hr_am, hr_pm),
    names_to = "measurement_time",
    values_to = "heart_rates"
)
```

```{r}
income <- tibble(religion = c("Agnostic","Atheist","Buddhist", "Catholic"), 
  X10k = c(12,25,3,75), 
  X10k_to_20k = c(34,14,6,60),
  X20k_to_30k = c(45,22,10,110))
```

```{r}
income %>% 
  pivot_longer(
    col = c(X10k, X10k_to_20k, X20k_to_30k),
    names_to = "income_level",
    values_to = "count"
  )
```
Pivot wider takes the values in a cell and turn them into new variable names.

# Splitting a value in one cell across two columns
```{r}
messy_orders <- tibble( name = c("Joe Bloggs","Jane Smith","Penny Jones"), 
  order_info = c("XY7282/29-06-19","LO0182/15-03-19","AS6812/04-06-19"), 
  order_amount = c(58.00,125.00,75.25))
```

Separate something from one cell into two columns
```{r}
clean_orders <- messy_orders %>% 
  separate(
    col = order_info,
    into = c("order_id", "order_date"),
    sep = "/"
  )
```

Unite
```{r}
clean_orders %>% 
  unite(
    col = "order_info",
    c(order_id, order_date),
    sep =  "/"
  )
```

# Practice

```{r}
hospitals <- read_csv("hospitals93to98.csv")
```

```{r}
hospitals %>% 
# Separate column IcdChapter
  separate(
    col = IcdChapter,
    into = c("icd_id", "icd_chapter"),
    sep = "\\."
  ) %>% 
# Pivot years to long format
  pivot_longer(
    cols = FY1993:FY1998,
    names_to = "year",
    values_to = "number_of_visits"
  ) %>% 
  
#Pivot field to wide format
  pivot_wider(
    names_from = Field,
    values_from = number_of_visits
  )
```

# Mini Lab Practice

```{r}
country_data <- read_csv("country_data.csv")
```


```{r}
country_data %>% 
# Deselect X1, a column which just includes the number of the row (as this will impact pivotting.)
  select(-X1) %>% 
# Seperate the number from the name of the country in the original column 'country'
  separate(
    col = country,
    into = c("country_number", "country"),
    sep = fixed("\\.")
  ) %>% 
# Pivot the multiple variables for year into one variable containing year.
  pivot_longer(
    cols = (Y1999 : Y2001),
    names_to = "year",
    values_to = "number"
  ) %>% 
# Pivot the 'type' column which currently contains two variables (populations, cases)
   pivot_wider(
    names_from = type,
    values_from = number
  ) %>%
# Unite the country code and country number to make one country code.
  unite(
    col = "country_code",
    c(code_letter, code_numb),
    sep = ""
  ) %>% 
  # remove 'country number' as it replicates data in our new country code column
  select(-country_number) %>% 
  # Rename our years with a numeric value
  mutate(
    year = case_when(
      year == "Y1999" ~ 1999,
      year == "Y2000" ~ 2000,
      year == "Y2001" ~ 2001
    )
  ) %>% view
```

