---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

# Why RegEx

helps us to make data more uniform. turn information written by humans into a single structure we can compare.
helps us to do text analysis
helps us to find things in data which are formatted or written in different ways.

They are case sensitive!
Checks for the pattern on every character

```{r}
single_string <- "string a"
strings <- c("string a", "string b", "string c", "string d", "striiiiiiing e")
```

## Single character
Strings containing letter a
```{r}
pattern <- "a"
```

```{r}
#Includes a -------------------------------------------------------------------
single_string %>% 
  str_detect(pattern)

strings %>% 
  str_detect(pattern)

# Does not include a ----------------------------------------------------------
single_string %>% 
  str_detect(pattern, negate = TRUE)
```

## Multi-character

Everything with the letters str
```{r}
pattern <- "str"

strings %>% 
  str_detect(pattern)
```

Everything containing any characters a, b or c or a -d or a-z and A-Z
```{r}
pattern <- "[abc]"

pattern <- "[a-d]"

pattern <- "[a-zA-Z]"

strings %>% 
  str_detect(pattern)
```

## Counting

```{r}
strings
```

Look for a pattern of multiple consecutive instances of i
```{r}
pattern <- "i{3}"

strings %>% 
  str_detect(pattern)
```

Look for any 3 consecutive letters - pattern <- "[a-z]{3}

Look for a pattern of 3 consecutive letters from a selection of characters, e.g. ngi finds nnn nng, nni, ing, iii, ggg, ggi, iin
```{r}
pattern <- "[ngi]{3}"

strings %>% 
  str_detect(pattern)
```

# Practice

Make a vector containing the following strings:
"We're having great fun learning about data analysis at CodeClan!"
"CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness"
"Our cohort code starts with DE, DG or D"
"You get a great view of the castle or the national conservatoire from just outside CodeClan's door"

```{r}
codeclan_vector <- str_c(c(
  "We're having great fun learning about data analysis at CodeClan!",
  "CodeClan is at 37 Castle Terrace in Edinburgh, ClockWise in Glasgow, and Wasps Academy in Inverness", 
  "Our cohort code starts with DE, DG or D", 
  "You get a great view of the castle or the national conservatoire from just outside CodeClan's door"))
```

Write and test regular expressions which will match:
Any string mentioning CodeClan
```{r}
codeclan_pattern <- "CodeClan"

codeclan_vector %>% 
  str_detect(pattern)
```

Any string containing two consecutive capital letters
```{r}
capital_pattern <- "[A-Z]{2}"

codeclan_vector %>% 
  str_detect(pattern2)
```

Any string not including the word "castle" or "Castle"
```{r}
castle_pattern <- "[cC]astle"

codeclan_vector %>% 
  str_detect(pattern3, negate = TRUE)
```

```{r}
strings <- c(
  "This string has an_address@email.com in it", 
  "This one has user.name@company.co.uk", 
  "Now we've got other_person_123@server.net and my.name@runningoutofideas.com"
)
```

# Detecting email address

Our strings
```{r}
email_pattern <- "[a-z]"

strings %>% 
  str_extract(email_pattern)
```

This has found the first match for each string.


```{r}
email_pattern <- "[a-z]@"

strings %>% 
  str_extract(email_pattern)
```

Finds the first match for this pattern in each string

```{r}
email_pattern <- "[a-z]*@"

strings %>% 
  str_extract(email_pattern)
```

Be careful with the star which means match any number of occurrences of the characters before. * matches 0 and up.

```{r}
email_pattern <- "[a-z]+@"

strings %>% 
  str_extract(email_pattern)
```

Plus means find at least one of the preceding characters

```{r}
email_pattern <- "[0-9a-z]+@[a-z]+"

strings %>% 
  str_extract(email_pattern)
```

```{r}
email_pattern <- "[0-9a-z._]+@[a-z.]+"

strings %>% 
  str_extract(email_pattern)
```

Still only returns one result per line

```{r}
email_pattern <- "[0-9a-z._]+@[a-z.]+"

strings %>% 
  str_extract_all(email_pattern)
```

```{r}
strings
email_pattern
```

Replace patterns in a string
```{r}
strings %>% 
  str_replace_all(email_pattern, "REDACTED")
```


# Practice
Make a vector containing the strings:
"Colin's mobile number definitely isn't 07777777777."
"A landline might be 01010 123456."
"Mhairi, Mandy, Colin and Steph have been some of your instructors so far."
"I'm called your_name_here and I live in your_town_here."

```{r}
numbers <- c(
  "Colin's mobile number definitely isn't 07777777777.",
  "A landline might be 01010 123456.",
  "Mhairi, Mandy, Colin and Steph have been some of your instructors so far.",
  "I'm called your_name_here and I live in your_town_here."
)
```

Use regular expressions to:
Extract all the names
```{r}
all_name_pattern <- "[A-Z][a-z]+"

numbers %>% 
  str_extract_all(all_name_pattern)
```

Extract the landline number
```{r}
landline_pattern <- "01[0-9 ]+"

numbers %>% 
  str_extract_all(landline_pattern)
```

```{r}
accurate_landline_pattern <- "01[0-9]{9}"

numbers %>% 
  str_replace_all(" ", "") %>% 
  str_extract_all(accurate_landline_pattern)
  
```


Insert your name and where you live into the fourth string

```{r}
your_name_pattern <- "your_name_here"
place_pattern <- "your_town_here"

numbers %>% 
  str_replace(your_name_pattern, "Lauren") %>% 
  str_replace(place_pattern, "Glasgow")
```

# Dealing with punctuation
A lot of characters have special meanings so we need to take care when dealing with punctuation

```{r}
punct_strings <- c(
  "Why is this a question?", 
  "I am a sentence with no punctation",
  "I am a well punctuated sentence.", 
  "WOW!"
)
```

Finds all punctuation
```{r}
punct_strings %>% 
  str_extract_all("[:punct:]")
```

Find 3 capital letter characters followed by any punctuation
```{r}
punct_strings %>% 
  str_extract_all("[A-Z]{3}[:punct:]")
```

```{r}
short_strings <- c(
  "abc",
  "Abc",
  "abC",
  "s"
)
```

Looking for lower case at the start of the string.
```{r}
short_strings %>% 
  str_extract_all("^[a-z]")
```

Lower case at the back of the string
```{r}
short_strings %>% 
  str_extract_all("[a-z]$")
```

Combo

Only one character long
```{r}
short_strings %>% 
  str_extract_all("^[a-z]$")
```

Match any when first is lower case and last is lowercase
```{r}
short_strings %>% 
  str_extract_all("^[a-z]+$")
```

Starts with a lowercase, any number of lowercase letters over 1 and finishes with a lowercase letter 
```{r}
short_strings %>% 
  str_extract_all("^[a-z]+[a-z]$")
```



# RegEx in a dataframe


```{r}
student <- c("Abigail", "Bjorn", "Caitlin")
age <- c(25, 56, 40)
emails <- c("abigal12@company.co.uk", "bsmith@gmail.co.uk", "caitlin_1980")
student_info <- data.frame(student, age, emails)
student_info 
```

```{r}
student_info %>% 
  mutate(email_detection = str_detect(emails, "[a-z0-9]+@[a-z0-9.]+"))

student_info %>% 
  mutate(email_end = str_extract(emails, "@[a-z.]+"))
```

