# Forward selection
Add the next best variable one at a time

# Backward selection
Remove the least predictove variable one at a time.

# Best subset selection
Test all combinations of subsetted variables for the optimal model of each size.
i.e. find the best possible model with one variable, then the best possible model with 2 variables, etc.
The computing power increases exponentially with the number of predictors

# Leaps Package
```{r}
library(tidyverse)
library(CodeClanData)
library(leaps)
```

Problems with leaps::regsubsets()

It may include some, but not all, dummy variables.
It may include interactions between variables not included in the model.
Models can only be tested with penalised measures of fit.

# Build a regression model for the charges of a policy.

## Forward selection -
```{r}
regsubsets_forward <- regsubsets(charges ~ ., 
                                 data = insurance, 
                                 # up to 8 predictors (the full dataset)
                                 nvmax = 8, 
                                 method = "forward")

summary(regsubsets_forward)
```
```{r}
summary(regsubsets_forward)$which
```

Plot the penalised measure, such as adjusted r squard
```{r}
plot(regsubsets_forward, scale = "adjr2")
```
Highest adjr2 model contains age, bmi, children, smokeryes, regionsoutheasy and regionsouthwest

Plot BIC (more parsimonious than adjr2)
```{r}
plot(regsubsets_forward, scale = "bic")
```

Lowest BIC score is for age, children and smokeryes

Plot the R-squared value of the best model found at each number of variables
```{r}
plot(summary(regsubsets_forward)$rsq, type = "b")
```
We are not significantly increasing r squared beyond 3 variables.

```{r}
plot(summary(regsubsets_forward)$bic, type = "b")
```
Lowest BIC scores around 4 variables

## Backward Selection

```{r}
regsubsets_backward <- regsubsets(charges ~ .,
                                  data = insurance,
                                  method = "backward")

sum_regsubsets_backward <-summary(regsubsets_backward)
sum_regsubsets_backward
```

```{r}
sum_regsubsets_backward$which
```

```{r}
plot(sum_regsubsets_backward$adjr2, type = "b")
```

```{r}
regsubsets_exhaustive <- regsubsets(
  charges ~ .,
  data = insurance,
  nvmax = 8,
  method = "exhaustive"
)
sum_regsubsets_exhaustive <- summary(regsubsets_exhaustive)
sum_regsubsets_exhaustive
```
```{r}
plot(sum_regsubsets_exhaustive$adjr2, type = "b")
```

# We then need to calculate the significance of our variables to ensure that the variables included in the model are statistically significant.

Check the specification of the 6 predictor model
```{r}
sum_regsubsets_exhaustive$which[6, ]
```
At this level, we include some but not all regions

Check significance of region with ANOVA
```{r}
mod_without_region <- lm(charges ~ age + bmi + children + smoker, data = insurance)

mod_with_region <-lm(charges ~ age + bmi + children + smoker + region, data = insurance)

anova(mod_without_region, mod_with_region)
```
The model with region is not significantly better than the model with it so disregard region.

Diagnostics
```{r}
par(mfrow = c(2,2))
plot(mod_without_region)
```

Check the 4 predictor model
```{r}
sum_regsubsets_exhaustive$which[4, ]
```
age, bmi, children and smoker

```{r}
pred_model_4 <- lm(charges ~ age + bmi + children + smoker, data = insurance)

summary(pred_model_4)
```

```{r}
par(mfrow = c(2, 2))
plot(pred_model_4)
```
These do not look good. We may need to do variable transformation before we do subset selection.

# How to do best subset selection after variable transformation
See notes


