---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
```

```{r}
students_big
```

# Recoding

changing a single categorical variable

## IfElse

```{r}
students_big %>% 
  mutate(older_than_12 = if_else(ageyears>12, "Older than 12", "12 or Younger"))
```

```{r}
students_big %>%
  mutate(gender = if_else(gender == "M", "Male", "Female"))
```
# Practice
change R and L to right and left
change football/soccer into 'football

```{r}
students_big %>% 
  mutate(handed = if_else(handed == "R", "right-handed", "left-handed"))
```

```{r}
students_big %>% 
  summarise(unique(favourite_physical_activity))
```
If Else, with the else as do nothing

```{r}
students_big %>% 
  mutate(favourite_physical_activity = if_else(favourite_physical_activity == "Football/Soccer", "Football", favourite_physical_activity))
```

## Recode

```{r}
students_big %>% 
  mutate(travel_to_school = recode(travel_to_school, "Car" = "car", "Walk" = "walk", "Bus" = "bus"))
```
Another way to see a list of all values.

```{r}
students_big %>% 
  group_by(travel_to_school) %>% 
  summarise (n=n())
```

Then we can replace whatever we haven't described with a default value

```{r}
students_big %>% 
  mutate(travel_to_school = recode(travel_to_school, "Car" = "car", "Walk" = "walk", "Bus" = "bus", .default = "other"))
```
# Practice
Using the data below, do the following:
Recode the subject numbers to be: 001, 002, 003 and 004
Recode the sex column to be “Male” and “Female”
Recode the compl column to be T = “True”, F = “False”, and NA otherwise

```{r}
data <- read.table(header = TRUE, text="
 subject sex control cond1 cond2 compl
       1   M     7.9  12.3  10.7  T
       2   F     6.3  10.6  11.1  T
       3   F     9.5  13.1  13.8  N
       4   M    11.5  13.4  12.9  F
")
```

```{r}
data %>% 
  mutate(
    subject  = recode (subject, 
                     "1" = "001", 
                     "2" = "002", 
                     "3" = "003", 
                     "4" = "004",
                     ),
    sex = if_else
          (sex == "M", "Male", "Female")
    ,
    compl = recode(compl, 
                 "T" = TRUE, 
                 "F" = FALSE, 
                 .default = NA)
    )
```

## Format better and remember you can stack mutations

## Case_When

```{r}
students_big %>% 
  select(ageyears) %>% 
  mutate(
    age_divisible = case_when(
    ageyears %% 2 == 0 ~ "Even",
    ageyears %% 3 == 0 ~ "Divisible by 3",
    ageyears %% 5 == 0 ~ "Divisible by 5",
    TRUE ~ "Other number")
  )
```

Recode and Case when need the outputs to be all the same type

# Practice

Using case_when, make a new column estimated_travel_time which is:
Equal to 60 when your region is London.
If you aren’t from London, then:
10 when your travel_to_school is “Walk”.
20 when your travel_to_schoolis “Car”.
For everyone one else then it should equal to 30.

```{r}
students_big %>% 
  mutate(estimated_travel_time =
           case_when(
             region == "London" ~ 60,
             travel_to_school == "Walk" ~ 10,
             travel_to_school == "Car" ~ 20,
             TRUE ~ 30
           )
           )
```

