---
title: "Solving Errors Lab"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

In this lab you will be presented with several code chunks that produce errors. Your task is to try and solve the errors and get the code chunk not only running without an error, but doing what it should be doing. For each chunk you will get a short description of what it is meant to be doing. 

Remember to create a new notebook and copy over the contents of the .Rmd file and solve them in there, rather than changing the classnotes! You'll see the error messages when you try to run the chunks in your notebook - keep fixing them until you don't get an error (and it's accomplishing the goal).

Use what you have learned in the errors lesson to solve these errors. Remember: problem solving and trying to interpret the error messages should be your first approach. Where this does not cut it there are several options: Google, StackOverflow, rubber ducking, specific function help files, checking data types and dimensions.

**Note:** You may have to fix more than one thing per code chunk to escape the errors!

```{r}
library(tidyverse)
library(tibble)
library(janitor)
```


# Error time!

## Error 1
**Goal:** *Create `some_letters` which is a character vector containing the following letters: "E", "R", "R", "O" and "R".*
```{r}
some_letters <- c("E", "R", "R", "O", "R")

```
<br>

## Error 2

**Goal:** *Create `message` which is a vector containing the following strings: "does", "this", "work" and "?", then convert the vector to uppercase using `toupper`.*
```{r}
message <- c("does", "this", "work", "?")

toupper(message)
```

## Error 3

**Goal:** *Select columns `mpg`, `cyl`, and `disp` from `mtcars` and keep only rows where `mpg` is less than 20.*
```{r}
mtcars %>% 
  select(mpg, cyl, disp) %>%
  filter(mpg < 20)

```

<br>

## Error 4

**Goal:** *Select columns `Sepal.Length`, `Sepal.Width`, and `Species` and arrange by `Sepal Length`.*

Add column Sepal.Length to select function
Run janitor to fix col names
```{r, warning = F, message = F}

iris %>% 
  select(Sepal.Width, Species, Sepal.Length) %>% 
  arrange(Sepal.Length)


```
<br>

## Error 5


**Goal:** *Find the average tooth length for where the supplement is vitamin C.*

Change filter to ==
Invert filter and summarise
```{r}
ToothGrowth %>%
  filter(supp == "VC") %>% 
  summarise(avg_tooth_len_vc = mean(len))
  

```
<br>

## Error 6

**Goal:** *Create a vector of animals (cat, dog, rabbit, horse), then use the `sample` function to get 10 random animals from those categories.* [**Hint:** you probably want to take a look at the `sample` help file.]

quotation around dogs
add replace = TRUE to sample as we want a sample larger than our list

```{r}
animals <- c("cat", "dog", "rabbit", "horse")

sample(animals, 10, replace = TRUE)

```
<br>

## Error 7

**Goal:** *Create dog_names and cat_names, then use `intersect` to check which common pet names are shared by both dogs and cats.*

Cat names fix assignment operator
add cat_names to intersect function
remove capital from Intersect


```{r}

dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names <- c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")

intersect(dog_names, cat_names)

```

<br>

## Error 8

**Goal:** *Filter weights equal to or below 150 grams, then create a new column 'weight_kg' that contains the weight in kilograms.*

```{r}
names(chickwts)
```

move =< to <=
mutate takes = not ==
```{r}
chickwts %>% 
  filter(weight <= 150) %>% 
  mutate(weight_kg = weight * 0.001)
```

## Error 9

**Goal:** *Create a tibble with column `alphabet` which contains the English alphabet and column `numbers` which has a unique number for each letter. Then remove rows where the letter is A, B or C.*

Load tibble

load tibble
LETTERS not a function
call alphabet not letters

alphabet != c("a", "b", "c")
compares alphabet[1] with "a", alphabet[2] with "b", alphabet[3] with "c", alphabet[4] with "d", alphabet[5] with "e", etc.



```{r}
tibble(alphabet = LETTERS,
       numbers = 1:26) %>% 
  filter(alphabet != c("A", "B", "C"))

```

## Error 10

The chunk below creates the data for this error (don't change anything in it):
```{r}
# Don't change anything in this code chunk
# but you will need to include it in your notebooks

set.seed(999)

scent_ratings <- tibble(
  iD = rep(1:10, each = 5),
  SCENT = rep(c("New book", "Cut grass", "Rose oil", "Fresh linen", "Wet dog"), 10),
  rating = sample(c(0:5, NA), 50, replace = TRUE),
  PersonAge = rep(NA, 50)
)
```

**Goal:** *Clean the variable names and drop missing ratings from `scent_ratings` and calculate an average rating per scent.*

Load janitor


```{r}

scent_ratings <- scent_ratings %>% 
  clean_names()

scent_ratings %>% 
  select (-person_age) %>% 
  drop_na()  %>% 
  group_by(scent) %>% 
  summarise(mean = mean(rating))

```

## Error 11

If you've made it this far and you can't get enough of errors, we've got a challenge for you! Come up with your own code chunk that produces an error. Make it full of mistakes or hide a few subtle mistakes in it, it's up to you!
