---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)

sample_200 <- read_csv("data/telecomms_churn_200_sample.csv") %>% 
  clean_names()
```

# Calculating a 200 observation bootstrap sample
Each resample needs to be the same size as the sample.

```{r}
bootstrap_resample_200 <- sample_200 %>% 
  rep_sample_n(size = 200, replace = T, reps = 5000) %>% 
  summarise(mean_tenure = mean(tenure))
```

How close is the bootstrapped sampling distribution of mean(tenure) to the one shown above?
```{r}
pop <- read_csv("data/telecomms_churn.csv") %>% 
  clean_names() %>% 
  rep_sample_n(size = 200, replace = F, reps = 5000) %>%
  summarise(mean_tenure = mean(tenure))
```

Plot a histogram of the bootstrapped sampling distribution
```{r}
bootstrap_resample_200 %>% 
  ggplot() +
  aes(x = mean_tenure) +
  geom_histogram(col = "white")

pop %>% 
  ggplot() +
  aes(x = mean_tenure) +
  geom_histogram(col = "white")
```

Calculate the standard error in mean(tenure) and compare it with the previous value 1.72 [remember: this is just the sd() of the sampling distribution]
```{r}
bootstrap_resample_200 %>% 
  summarise(se_tenure = sd(mean_tenure))
```

Slightly higher standard error. 0.026 of a month, less than one day.

# Confidence interval from a bootstrap distribution

## Percentile Method
```{r}
ci_95 <- bootstrap_resample_200 %>% 
  summarise(mean_bootstrap = mean(mean_tenure),
            lower_bound = quantile(mean_tenure, 0.025),
            upper_bound = quantile(mean_tenure, 0.975)
  )
ci_95
```

## 99% Confidence
```{r}
ci_99 <- bootstrap_resample_200 %>% 
  summarise(mean_bootstrap = mean(mean_tenure),
            lower_bound = quantile(mean_tenure, 0.005),
            upper_bound = quantile(mean_tenure, 0.995)
  )
ci_99
```
This gives us a wider range, from 28.17 - 37.11 months, which we can be 99% confident that the population value falls within.
It is a bigger interval that we are more confident in.

# Infer Package Version of this workflow

```{r}
infer_resample <- sample_200 %>% 
  specify(response = tenure) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean") 

infer_ci <- infer_resample %>% 
  get_ci(level = 0.95, type = "percentile")
```

# visualise our confidence interval
```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_ci(endpoints = infer_ci)
```
# Find the mean using regular dplyr
```{r}
infer_resample %>% 
  summarise(mean = mean(stat))
```

# Task
- same workflow with 300 (400, 500) observation sample
- mean tenure as a sample distribution and point estimate
- 95% CI - compare with ci_200

```{r}
sample_300 <- read_csv("data/telecomms_churn_300_sample.csv") %>% 
  clean_names()

bootstrap_300 <- sample_300 %>% 
  specify(response = tenure) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

ci_bootstrap_300 <- bootstrap_300 %>% 
  get_ci(level = 0.95, type = "percentile")

ci_bootstrap_300
infer_ci
```

with a larger dataset, my confidence interval gets narrower. Because we can be more confident in our results, we can achieve higher levels of certainty on a smaller interval.

```{r}
sample_400 <- read_csv("data/telecomms_churn_400_sample.csv") %>% 
  clean_names()

bootstrap_400 <- sample_400 %>% 
  specify(response = tenure) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

ci_bootstrap_400 <- bootstrap_400 %>% 
  get_ci(level = 0.95, type = "percentile")

ci_bootstrap_400

ci_bootstrap_300

infer_ci
```

```{r}
sample_500 <- read_csv("data/telecomms_churn_500_sample.csv") %>% 
  clean_names()

bootstrap_500 <- sample_500 %>% 
  specify(response = tenure) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

ci_bootstrap_500 <- bootstrap_500 %>% 
  get_ci(level = 0.95, type = "percentile")

ci_bootstrap_500
ci_bootstrap_400
ci_bootstrap_300
infer_ci
```


# Confidence Intervals for a proportion

```{r}
bootstrap_churn <- sample_200 %>% 
  specify(response = churn, success = "Yes") %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "prop")

churn_ci <- bootstrap_churn %>% 
  get_ci(level = 0.95, type = "percentile")

churn_ci
```

```{r}
bootstrap_churn %>% 
  visualise(bins = 30) +s
  shade_ci(endpoints = churn_ci)
```

