---
title: "R Notebook"
output: html_notebook
---

```{r}
library(stringr)
```


# Writing Functions

Pseudo
```{r}
function_name <- function(argument_1, argument_2, argument_3 = value) {
  
  do something with arguments
  
  return(output)
}
```

multiply by 5
```{r}
multiply_by_5 <- function(x) {
  return(x* 5)
}
```

multiply by each other
```{r}
multiply_by <- function(x, y) {
  x*y
}
```

function with a default
```{r}
multiply_by_5_unless <- function(x, y=5) {
  return(x*y)
}
```

# Practice
Write a function make_price() that takes as input any whole number, and returns it written as a price (i.e. with a ‘£’ in front of the number). For example 5 would become £5, and 123 would become £123. Try out your function a few times to test that it works.

```{r}
make_price <- function(x) {
  return(paste("£", x, sep = ""))
}

make_price(199)
```

Write a function make_fancy_price() that takes any decimal number and returns it as a ‘written out’ price in the form “# pounds # pence”. If the input number corresponds only to whole pounds or whole pence, then return only that part of the price (e.g. “# pounds” or “# pence”). You’ll also need to round to the nearest number of pence.

```{r}
# if only whole numbers return "x pounds"
# if over 1 and not whole return "x pounds, x pence"
# if less than 1, return "x pence"

after_decimal <- function(x) {
  while (x > 0.99) {
    x <- x-1
  }
  x <- format(x, nsmall = 2)
  x <- (as.character(x))
  x <- str_sub(x, -2, -1)
  return(x)
}

undebug(after_decimal)

after_decimal(8.99)
y <- after_decimal(9.80)

y
class(y)
length(y)
str_sub(y, -2, -1)
```


```{r}
before_decimal <- function(x) {
  while (x != round(x)) {
    x <- x - 0.01
  }
  return(x)
}

before_decimal(8.01)
```

```{r}
# before_decimal_efficient <- function(x) {
#   while (x != round(x)) {
#     while (x %% .1 != 0) {
#       x <- x - 0.01
#     }
#     x <- x - 0.1
#   }
# }
# 
# before_decimal_efficient(8.20)
```

```{r}
make_fancy_price <- function(x) {
  if (x < 1) {
    return(paste(after_decimal(x), "pence"))
  }
  if(x > 1 & x %% 1 == 0) {
    return(paste(before_decimal(x), "pounds"))
  } else {
    return(paste(before_decimal(x), "pounds", after_decimal, "pence"))
  }
}

make_fancy_price(9.68)

```


Here’s some examples of inputs and expected outputs.

Input	Output
3.23	3 pounds 23 pence
4.5	4 pounds 50 pence
8	8 pounds
50.012	50 pounds 1 pence
0.546	55 pence
You may need to think about the DRY principle.
Perhaps think about separate functions to format pounds and pence correctly.
Remember you can call one function from inside another.

