---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
# Wine

```{r}
library(tidyverse)
library(janitor)
```


## Read in the data and investigate it

```{r}
wine_ratings <- read_csv("data/wine_ratings.csv")

# Count NAs in multiple columns

wine_ratings %>% 
  summarise(
    nadescription = sum(is.na(description)),
    nacountry = sum(is.na(country))
  )

# Count NAs in all columns quick
wine_ratings %>% 
  is.na() %>% 
  colSums()

# Count NAs in all columns pretty
wine_ratings %>% 
  summarise(across(.fns = ~sum(is.na(.x))))

# Variable names
colnames(wine_ratings)

# How many values are in each column?
dim(distinct(wine_ratings, designation))

dim(distinct(wine_ratings, region_1)) %>% 
  class()

wine_ratings %>% 
  summarise(across(where(is.character), .fns = ~length(unique(.x))))
```



# Select columns title, taster_name, points, price, and country, then drop all NAs. Save into a variable called wine_subset. Use this variable for all subsequent tasks.

```{r}
wine_subset <- wine_ratings %>% 
  select(title, taster_name, points, price, country) %>% 
  drop_na()
```



# Find the 5 countries with the highest mean points.

```{r}
highest_mean_points <- wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_points = mean(points)) %>% 
  arrange(desc(mean_points)) %>% 
  head(5)
```


# Find the 5 countries with the lowest mean prices.

```{r}
wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_price = mean(price)) %>% 
  arrange(mean_price) %>% 
  head(5)
```
 # Compare the 5 countries with the highest mean points with the 5 countries with the highest median points
 
 Our first 4 countries stay the same by mean or by median, but Switzerland has a higher mean than Australia and Australia has a higher median than Switzerland

```{r}
highest_mean_points

highest_median_points <- wine_subset %>% 
  group_by(country) %>% 
  summarise(median_points = median(points)) %>% 
  arrange(desc(median_points)) %>% 
  head(5)

class(highest_mean_points)

highest_mean_country <- highest_mean_points %>% 
  select(country)

highest_median_country <- highest_median_points %>% 
  select(country)

highest_mean_country == highest_median_country
```

# Compare the 5 countries with the lowest mean prices with the the 5 countries with the lowest median prices

Our 4th and 5th lowest priced countries change depending on whether we calculate by mean or median.

```{r}
lowest_mean_countries <- wine_subset %>% 
  group_by(country) %>% 
  summarise(mean_price = mean(price)) %>% 
  arrange(mean_price) %>% 
  head(5) %>% 
  pull(country)

lowest_median_countries <- wine_subset %>% 
  group_by(country) %>% 
  summarise(median_price = median(price)) %>% 
  arrange(median_price) %>% 
  head(5) %>% 
  pull(country)

lowest_mean_countries == lowest_median_countries
```

# Find the names of all the tasters using distinct. Then take a random sample of 5 names using sample_n. Save these 5 names into a new variable called taster_sample which should be a character vector containing only the names.

```{r}
taster_sample <- wine_subset %>% 
  distinct(taster_name) %>% 
  slice_sample(n = 5) %>% 
  pull(taster_name)

class(taster_sample)

taster_sample
```


# Create a character vector called country_sample using distinct and pull, but this time use sample_frac to pick 10 countries at random.

```{r}
country_sample <- wine_subset %>% 
  slice_sample(prop = .001) %>% 
  pull(country)

country_sample
    
```

# Filter the rows of wine_subset so you are left only with data from the 5 tasters in your taster_sample and from the countries in your country_sample.

```{r}
wine_sample <- wine_subset %>% 
  filter(
    country %in% country_sample &
      taster_name %in% taster_sample
  )
dim(wine_sample)
```


Question 9

Using your wine_sample, create a new column difference_from_avg which contains the difference between each wine’s rating (points) and the average rating for that taster, then arrange the differences to see the highest positive difference first. [Hint: you will first have to calculate the average points for each taster]

```{r}
wine_sample %>% 
  group_by(taster_name) %>% 
  mutate(taster_avg = mean(points)) %>% 
  ungroup() %>% 
  mutate(difference_from_avg = points - taster_avg) %>% 
  select(-taster_avg) %>% 
  arrange(desc(difference_from_avg))

```

