---
title: "R Notebook"
output: html_notebook
---

# Single Step Experiment
e.g. one roll of our two die


```{r}
# set up the sample and event spaces again
library(prob)
library(tidyverse)

# get all possible combos of 2 die rolls
s <- rolldie(2)
size_s <- nrow(s)
size_s
```

```{r}
# find when the two separate die rolls total 7
total_seven <- s %>%
  dplyr::filter(X1 + X2 == 7)

# how many combos exist
size_total_seven <- nrow(total_seven)
size_total_seven
```

```{r}
# using this to find a probability

prob_total_seven <- size_total_seven / size_s
prob_total_seven
```

# Multiplie step experiments
e.g. throw one die, see the result, then flip a coin

```{r}
# set up one roll of a dice
s_die <- rolldie(1)

# set up one toin coss
s_coin <- tosscoin(1)

# merge them together
# The merge() function takes a row from s_die and associated it with each row in s_coin.
s_total <- merge(s_die, s_coin)
s_total
```

# Permutations 

```{r}
library(gtools)
```

```{r}
# create vector of choices contains ten employees, n = 10. 
employees <- c("Alice", "Bob", "Carole", "David", "Elaine", "Fred", "Gabra", "Hua", "Inigo", "Hugo")

# Choose three at a time, r = 3
# Don't replace an employee once chosen: repeats.allowed = FALSE
seniors <- as_tibble(
  permutations(
    # n = number of options
    n = 10, 
    # r = how many people we want in our outcome
    r = 3, 
    # v = vector of options
    v = employees, 
    # more than one employee can't fill the same position therefore no repetition
    repeats.allowed = FALSE
  )
)
```

```{r}
# rename the variables in seniors to something better
seniors <- seniors %>%
  rename(
    CEO = V1,
    CTO = V2,
    COO = V3
  )

# let's see what the seniors matrix looks like
seniors
```

What is the probability of Alice being selected as CEO and Bob being selected as CTO, if the selection process is random?

```{r}
# first get the event space
alice_CEO_bob_CTO <- seniors %>%
  filter(CEO == "Alice", CTO == "Bob")
alice_CEO_bob_CTO
```
```{r}
prob_alice_CEO_bob_CTO <- nrow(alice_CEO_bob_CTO) / nrow(seniors)
prob_alice_CEO_bob_CTO
```

# Task

Task - 10 mins A licence plate begins with three letters from a possible set of A, B, C, D, E. How many permutations are possible if repetition of letters is:

not allowed [Hint - how many letters are possible for first position, how many for second if one is ‘used up’, how many for third if two are ‘used up’?]


```{r}
# variable with the possibile letters
poss_letters <- c("A", "B", "C", "D", "E")

# find permutations with repeat = false

poss_licence_plate_no_rep <- as_tibble(
  permutations(
    n = length(poss_letters),
    # how many letters we want
    r = 3,
    v = poss_letters,
    repeats.allowed = FALSE
  )
)
nrow(poss_licence_plate_no_rep)
```


allowed [Hint - all that changes from last answer is that letters don’t get ‘used up’]

```{r}
poss_licence_plate_with_rep <-
  as_tibble(
    permutations(
      n = length(poss_letters),
      r = 3,
      v = poss_letters,
      repeats.allowed = TRUE
    )
  )
nrow(poss_licence_plate_with_rep)
```

Harder Task - 10 mins A random four digit pin is set on a gym lock. What is the probability that no digits are repeated?
```{r}
num <- c(0:9)

# Define event space for locker combo
s_locker <- as_tibble(
  permutations(
    n = length(num),
    r = 4,
    v = num,
    repeats.allowed = TRUE
  )
)

# Define event space for locker combo with no repeats
s_locker_no_rep <- as_tibble(
  permutations(
    n = length(num),
    r = 4,
    v = num,
    repeats.allowed = FALSE
  )
)

prob_locker_no_rep <- nrow(s_locker_no_rep) / nrow(s_locker)
prob_locker_no_rep

```

It is as likely to contain repeats as not.

# Combinations

```{r}
# Set up deck
cards <- cards(jokers = FALSE)

cards <- cards %>% 
  mutate(card = str_c(rank, str_sub(suit, start = 1, end = 1))) %>%
  pull(card)
```

```{r}
# find the number of combinations for a 5 card hand

five_card_hands <-
  combinations(
    n = 52,
    r = 5,
    v = cards,
    repeats.allowed = FALSE
  )

nrow(five_card_hands)
```

If all we are interested in is the number of combinations(), the choose() function gives us this much more quickly:

```{r}
num_5_card_hands <- choose(52, 5)
num_5_card_hands
```

# Task
Given a lottery with 50 numbers drawn in lots of 6, how many possible winning combinations are there?
If you buy 10 tickets, each with a different set of numbers, what is your probability of winning?
Remember, order doesn’t matter in lottery draws, and no repetition is allowed (once a numbered ball pops out the machine, it isn’t put back in).

```{r}
lottery_num <- c(1:50)

s_lottery_draw <- choose(50, 6)
```

```{r}
prob_win_ten_tickets <- 
  10/s_lottery_draw
```

# Task
I have five flavours of ice cream to choose from and a bowl that can hold three scoops. How many different variations in bowls of ice cream can I have assuming:


```{r}
flavours <- c("vanilla", "chocolate", "raspberry", "strawberry", "pistachio")
```

##Each scoop must be a different flavour

```{r}
choose(5, 3)
```


## I can repeat flavours if I wish?

```{r}
nrow(
  as_tibble(
    combinations(n = 5, r = 3, v = flavours, repeats.allowed = T)
  )
)
```

