---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)

telco <- read_csv("data/telecomms_churn.csv") %>% 
  clean_names()

head(telco)
```

# Our imaginary population

```{r}
summary_pop <- telco %>% 
  summarise(
    mean_monthly_charges = mean(monthly_charges),
    mean_tenure = mean(tenure),
    prop_churn = mean(churn == "Yes")
  )

summary_pop
```

```{r}
telco %>% 
  ggplot() +
  aes(x = monthly_charges) +
  geom_histogram(col = "white")
```

```{r}
telco %>% 
  ggplot() +
  aes(x = tenure) +
  geom_histogram(col = "white")
```

```{r}
telco %>% 
  ggplot(aes(x = churn)) +
  geom_
```

# Sample
## Simple Random Sampling
every observation in the population has an equal chance of being selected

```{r}
library(infer)

# take 200 observations from the data, one time.
sample_200 <- telco %>% 
  rep_sample_n(size = 200, reps = 1, replace = F)
sample_200

# our data comes out grouped by replicate
groups(sample_200)
```

```{r}
summary_sample_200 <- sample_200 %>% 
  ungroup() %>% 
  summarise(
    mean_monthly_charges = mean(monthly_charges),
    mean_tenure = mean(tenure),
    prop_churn = mean(churn == "Yes")
  )

# These are statistics or point estimates
summary_sample_200

# These are parameters
summary_pop
```

# Sampling error

## In theory
sampling error = point estimate - population parameter
 
```{r}
summary_sample_200 - summary_pop
```
 
# Resampling
A way of imitating the act of collecting lots of different samples
 
```{r}
rep_sample_200 <- telco %>%
  rep_sample_n(size = 200, reps = 5000) %>%
  summarise(
    mean_monthly_charges = mean(monthly_charges), 
    mean_tenure = mean(tenure),
    prop_churn = mean(churn == "Yes")
  ) 
rep_sample_200
```
# Sample distribution
```{r}
# here ..density.. tells ggplot to use the probability density rather than count
# in the histogram
monthly_charges_plot <- rep_sample_200 %>%
  ggplot(aes(x = mean_monthly_charges)) + 
  geom_histogram(col = "white", fill = "steel blue", alpha = 0.7) +
  labs(x = "mean monthly_charges from each sample")
monthly_charges_plot
```
 
# Standard Error

Standard error is the standard deviation of the sampling distribution.

```{r}
std_errors <- rep_sample_200 %>% 
  summarise(
    se_month_charges = sd(mean_monthly_charges),
    se_tenure = sd(mean_tenure),
    se_prop_churn = sd(prop_churn)
  )
std_errors
```

```{r}
library(fastGraph)
```

## What proportion of our monthly charges fall between 60 and 70?

```{r}
shadeDist(
  xshade = c(60, 70), 
  lower.tail = FALSE, 
  ddist = "dnorm", 
  parm1 = mean(rep_sample_200$mean_monthly_charges), 
  parm2 = std_errors$se_month_charges, 
  xlab = "mean_monthly_charges",
  col = c("black", "grey")
)
```
98% of the samples we have pulled have a mean between 60 and 70.
Following our simulation, 98% of our samples fall that they have a mean between 60 and 70.
This is not inferential statistics yet.

This lesson shows us that sample distributions can help us to model what is happening in a population

